\chapter{Number-Theoretic Algorithms}
\section{GCD and Extended Euclidean Algorithm}

\subsection{GCD}
The greatest common divisor (GCD) of two integers
$a$ and $b$ is the largest integer that divides both $a$ and $b$.
While seemingly simple, the GCD is a fundamental concept in number theory
and used extensively when working with integers over finite fields.


\begin{lstlisting}[language=Python,
  caption={GCD algorithm in python}
]

def gcd(a, b):
    while b != 0:
        a, b = b, a%b

    return a
 \end{lstlisting}

This algorithm relies on the following theorem, allowing us to work with the remainder (until it is zero).
\begin{theorem}
  Consider $a=b\cdot q +r$ then $GCD(a,b)= GCD(b,r)$.
\end{theorem}

\begin{proof}
  Consider $a=b\cdot q +r$, and $GCD(a,b)=x$.

  $x$ divides both $a$ and $b$ by definition. So $x$ also divides 
  $b\cdot q+ r$ (otherwise it wouldn't be the GCD of $a$ and $b$).
  Thus, $x$ must divide $r$ as well. 

  Let $GCD(b,r)=y$, Since $x$ also divides $r,b$ we know that $y\ge x$.
  Suppose $r=a-b\cdot q=b\cdot(-q) +a$ (i.e., re-ordering the equation). 
  That is, $y$ is the GCD of $r,b$ so it must also divide $b\cdot(-q) +a$: 
  $y$ is a common divisor of $a,b$ and from $GCD(a,b)=x\ge y$.
  
  Since $x\ge y \ge x$ they must be equal.
\end{proof}


\subsection{Extended Euclidean Algorithm (inverse algorithm)}
\label{sec:finite-field:extended-euclidean}
The Extended Euclidean Algorithm is an important algorithm in any number theory course.
This algorithm is used to find the inverse of an element in a finite field.
\paragraph{Bezout's identity}
Bézout's identity states that for any two integers $a$ and $b$ (not both zero), there exist integers $x$ and $y$ such that:
$$ ax + by = \gcd(a, b) $$
These integers $x$ and $y$ are called Bézout coefficients for $(a, b)$. Furthermore, $\gcd(a, b)$ is the smallest positive integer that can be expressed in the form $ax + by$.

The following proof is an iteration of the above explanation.
\begin{proof}
    Let $S = \{ax + by \mid x, y \in \mathbb{Z}, ax + by > 0\}$. By the Well-Ordering Principle, $S$ has a smallest element $d$.
    We claim $d = \gcd(a, b)$.
    First, $d$ must divide $a$. If not, $a = qd + r$ with $0 < r < d$.
    But $r = a - qd$ is a linear combination of $a$ and $b$ smaller than $d$, contradicting minimality. Thus $d \mid a$ and similarly $d \mid b$.
    Since $d$ is a common divisor, $d \le \gcd(a, b)$. Also, $\gcd(a, b)$ divides any linear combination of $a$ and $b$, so $\gcd(a, b) \mid d$, implying $\gcd(a, b) \le d$.
    Therefore, $d = \gcd(a, b)$.
\end{proof}

\paragraph{Using Bezout to Find Inverse in Field}
As a result of Bezout's identity, we can find the inverse of an element in a finite field $GF(q)$. 
in short, in a finite field $q$ is a prime right? so $GCD(a,q)=1$ for every $a\in GF(q)$.
As a result, we have $ax+qy=ax+0=GCD(a,q)=1$ (where $qy=0$ since we are in $GF(q)$), and thus $ax=1 \mod q$. and luckily, the 
extended Euclidean algorithm can find this $x$ and this $y$ (we toss the $y$ when we just want the inverse of $a$ in $GF(q)$).

The same idea works in rings: if $GCD(a,q)=1$ then there exists an inverse.
if $GCD(a,q)\neq 1$, there isn't any inverse. 


\paragraph{Algorithm}

The algorithm is basically applying the GCD algorithm, but keeping track of the coefficients $x$ and $y$ in the process.

\begin{lstlisting}[language=Python,
  caption={Extended Euclidean Algorithm in python}
]
def extended_gcd(a, b):
    # Base case: gcd(a, 0) = a, and a = a*1 + 0*0, so x=1, y=0.
    if b == 0:
        return a, 1, 0

    # Recursive call returns (g, x', y') such that g = b*x' + r*y',
    # where r = a mod b. We'll rename them to x, y for brevity.
    g, x, y = extended_gcd(b, a % b)

    # Now convert coefficients for (b, r) into coefficients for (a, b).
    # Since r = a - (a // b) * b, substitute into g = b*x + r*y:
    #   g = b*x + (a - (a//b)*b) * y
    #     = a*y + b*(x - (a//b)*y)
    # So the new x is the old y, and the new y is x - (a//b)*y.
    return g, y, x - (a // b) * y
\end{lstlisting}

\red{below paragraphs are generated via gemini3, i've read through them though.}

\paragraph{Matrix Intuition}
Let us find the iterative procedure (which is also important for the Half-GCD algorithm).

We can view the Euclidean algorithm steps as matrix transformations.
We start with the identity relations:
\[
\begin{pmatrix} a \\ b \end{pmatrix} = \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix} \begin{pmatrix} a \\ b \end{pmatrix}
\]
In each step of the Euclidean algorithm, we perform the division $a = q \cdot b + r$, which implies $r = a - q \cdot b$.
If we want to transition from the state $(a, b)$ to $(b, r)$, we can apply the following transformation matrix:
\[
\begin{pmatrix} b \\ r \end{pmatrix} = \begin{pmatrix} 0 & 1 \\ 1 & -q \end{pmatrix} \begin{pmatrix} a \\ b \end{pmatrix}
\]

We found a transition matrix right? let's call it $M$. what happens if we apply $M$. This matrix essentially help us transofrm from $GCD(a,b)$ state to the next $GCD(b,r)$ state.

Similarly, if we apply this matrix again on the $(b,r)$ vector, we'll get the next GCD state.

So lets us apply this state again and again until we reach the state $GCD(g,0)$. Namely, we reached our goal and found the GCD.

$\begin{pmatrix}gcd \\0 \end{pmatrix} = M \cdot M \cdot M \cdots M \cdot \begin{pmatrix} a \\ b \end{pmatrix}$

What is of all these $M$ matrices? it's a matrix too: 
$M^* = M \cdot M \cdot M \cdots M = \begin{pmatrix} M_{11}^*   & M_{12}^* & \\ M_{21}^* & M_{22}^* \end{pmatrix} $

Noitce that 
$\begin{pmatrix}gcd \\0 \end{pmatrix} = \begin{pmatrix} M_{11}^*   & M_{12}^* & \\ M_{21}^* & M_{22}^* \end{pmatrix} \cdot \begin{pmatrix} a \\ b \end{pmatrix}$

That is, we have $gcd=M_{11}^*a + M_{12}^*b$. That is, if we stored our matrix and kept it on the side, we can easily grab $x$ and $y$ such that $gcd=a\cdot x+ y\cdot b$.


% \paragraph{Invariant Explanation}
% To make it even clearer, let's denote the initial inputs as $A$ and $B$.
% The variables $x_0, y_0$ track the coefficients for the current value of variable \texttt{a}, and $x_1, y_1$ track the coefficients for the current value of variable \texttt{b}.
% Throughout the execution of the loop, the following invariant holds:
% \[
% \begin{pmatrix} \texttt{a} \\ \texttt{b} \end{pmatrix} = \begin{pmatrix} x_0 & y_0 \\ x_1 & y_1 \end{pmatrix} \begin{pmatrix} A \\ B \end{pmatrix}
% \]
% This means that at any step:
% \[
% \texttt{a} = x_0 \cdot A + y_0 \cdot B \quad \text{and} \quad \texttt{b} = x_1 \cdot A + y_1 \cdot B
% \]
% When the loop terminates ($\texttt{b}=0$), the variable \texttt{a} holds the GCD, and the invariant tells us that $\text{GCD}(A, B) = x_0 \cdot A + y_0 \cdot B$.
% Thus, the returned $x_0$ and $y_0$ are the Bezout coefficients.

% \paragraph{Deriving the Update Rules}
% The code updates \texttt{x0, x1} and \texttt{y0, y1} by performing a matrix multiplication.
% Let $M = \begin{pmatrix} x_0 & y_0 \\ x_1 & y_1 \end{pmatrix}$ be the matrix storing the current coefficients.
% In each step of the loop, we apply the transformation $T = \begin{pmatrix} 0 & 1 \\ 1 & -q \end{pmatrix}$.
% The new coefficient matrix $M_{new}$ is the product $T \times M$:
% \[
% M_{new} = \begin{pmatrix} 0 & 1 \\ 1 & -q \end{pmatrix} \begin{pmatrix} x_0 & y_0 \\ x_1 & y_1 \end{pmatrix} 
% = \begin{pmatrix} x_1 & y_1 \\ x_0 - q \cdot x_1 & y_0 - q \cdot y_1 \end{pmatrix}
% \]
% This matrix multiplication exactly matches the Python tuple assignment:
% \begin{itemize}
%     \item The first row of $M_{new}$ becomes the new $(x_0, y_0)$, which takes values from the old second row $(x_1, y_1)$.
%     \item The second row of $M_{new}$ becomes the new $(x_1, y_1)$, calculated as $(x_0 - q x_1, y_0 - q y_1)$.
% \end{itemize}

\section{Chinese Remainder Theorem}
\label{sec:crt}

The Chinese Remainder Theorem (CRT) addresses the following question: given a system of congruences
\[
  x \equiv a_1 \pmod{n_1}, \quad x \equiv a_2 \pmod{n_2}, \quad \ldots, \quad x \equiv a_k \pmod{n_k},
\]
does a solution $x$ exist, and if so, is it unique?

\begin{theorem}[Chinese Remainder Theorem]
  Let $n_1, n_2, \ldots, n_k$ be pairwise coprime positive integers, and let $a_1, a_2, \ldots, a_k$ be arbitrary integers.
  Then there exists an integer $x$ satisfying all congruences simultaneously, and it is unique modulo $N = n_1 \cdot n_2 \cdots n_k$.
\end{theorem}
(Without proof)

The solution is computed constructively. Define $N = \prod_{i=1}^k n_i$ and $N_i = N / n_i$.
Since $\gcd(N_i, n_i) = 1$, the inverse $M_i = N_i^{-1} \pmod{n_i}$ exists and can be found via the Extended Euclidean Algorithm.
The unique solution is then $x = \sum_{i=1}^k a_i N_i M_i \pmod{N}$.

If $n_1,n_2,\dots,n_k$ are not pairwise coprime then it is possible no solution exists, or it isn't unique. 

\begin{lstlisting}[language=Python, caption={Chinese Remainder Theorem solver}]
from math import prod

def mod_inverse(a, n):
    # Inverse of a modulo n via the Extended Euclidean Algorithm
    _, x, _ = extended_gcd(a, n)  
    return x % n

def crt(remainders, moduli):
    # remainders: list [a_1, ..., a_k]
    # moduli:     list [n_1, ..., n_k], must be pairwise coprime

    N = prod(moduli)          # N = n_1 * n_2 * ... * n_k

    x = 0
    for a_i, n_i in zip(remainders, moduli):
        N_i = N // n_i               # N_i = N / n_i
        M_i = mod_inverse(N_i, n_i)  # M_i = N_i^{-1} mod n_i
        x += a_i * N_i * M_i

    return x % N              # unique solution mod N
\end{lstlisting}

\section{Generator}
We've specified in \autoref{theorem:num-generators} that we have $\phi(q-1)$ generators in a finite field $\mathbb{F}_q$ (notice we don't have $\phi(q)$ generators).

So how do we find them?
Let us sample $g \leftarrow \mathbb{F}_q^*$. We will test if it is a primitive element.

By \autoref{theorem:lagrange-groups}, the order of any element in $\mathbb{F}_q^*$ divides $q-1$.
For each prime divisor $p_i$ of $q-1$, we check whether $g^{(q-1)/p_i} = 1$.
If so, $g$ is not a generator: the order of $g$ divides $(q-1)/p_i < q-1$, so it cannot equal $q-1$.


\begin{insight}[Why Does it Work?]
It suffices to check only the prime divisors of $q-1$, rather than all divisors of $q-1$.

\textbf{If $g$ is not a generator, some check will catch it.}
Suppose $\mathrm{ord}(g) = d$ where $d \mid (q-1)$ and $d \neq q-1$.
Since $d$ is a proper divisor of $q-1$, the quotient $(q-1)/d$ is an integer greater than or equal to $2$, so it has at least one prime factor; call it $p_j$.

Because $p_j \mid (q-1)/d$, we can write $(q-1)/d = p_j \cdot m$ for some integer $m$, so $q - 1 = d \cdot p_j \cdot m$.
In particular, $d \mid (q-1)/p_j$.
Therefore:
$$g^{(q-1)/p_j} = \left(g^{d}\right)^{(q-1)/(d \cdot p_j)} = 1^{(q-1)/(d \cdot p_j)} = 1.$$
The check at the prime $p_j$ evaluates to $1$, so $g$ is correctly rejected.

\textbf{If every check passes, $g$ is a generator.}
Suppose $g^{(q-1)/p_i} \neq 1$ for every prime divisor $p_i$ of $q-1$.
This means $\mathrm{ord}(g)$ does not divide $(q-1)/p_i$ for any $p_i$.
The only divisor of $q-1$ with this property is $q-1$ itself, so $\mathrm{ord}(g) = q-1$ and $g$ is a generator.
\end{insight}

\section{Roots of Unity}

\chapter{Polynomial Algorithms}\label{chapter:polynomial-algorithms}

\section{Polynomial evaluation}
Polynomial evaluation is the process of computing the value of a polynomial at a given point.
This is a fundamental operation in many algorithms and applications. 
There are several methods to evaluate polynomials, with varying levels of efficiency.

I highly recommend reading the section on polynomial evaluation before 
proceeding (refer to~\ref{sec:poly-eval}), as I have personally used it in 
practice to both verify the correctness of my algorithms (as outlined
in paper~\cite{slicedPIR} and gain a deeper 
understanding of the capabilities and limitations of polynomials.
% To see the formal definition of a polynomial, refer to \autoref{def:poly-eval}

\subsection{Evaluation Algorithms}
Evaluating a polynomial at a point is a straightforward process.
Substituting the point for $X$ in the polynomial gives the result. 
However, Horner's rule offers a more efficient method.

Horner's rule reduces the number of multiplications required to evaluate the polynomial.
It rewrites the polynomial in a nested form, making computation more efficient.

To apply Horner's rule to evaluate a polynomial $P(x) = a_0 + a_1x + a_2x^2 + \dots + a_nx^n$ at a 
point $x = r$, follow these steps:

\begin{lstlisting}[language=Python,
  caption={Horner's rule algorithm in python}
]
def horner(p: List[int], r: int) -> int:
    result = p[-1]
    for i in range(len(p) - 2, -1, -1): # iterate from n-1 to 0
        result = result * r + p[i] # result = ((((....)x+a_3)x+a_2)x+a_1)x +a_0 = p(r)
    return result
\end{lstlisting}\label{algorithm:polynomial:evaluation}


\section{Coefficient vs Point-Value representation}\label{polynomials:point-val-representation}
We've defined one representation for polynomials, the coefficient representation~\ref{polynomials:coeffcient-pres},
But there is another representation, which is used in numerous algorithms (e.g.,~\ref{schawtz-zippel},\ref{polynomial:multiplication}):

\begin{definition}
The Point-Value representation of a polynomial $P(x)$ of degree-bound $n$ is a set of $n+1$ point-value pairs. 
\(\{ (x_0,P(x_0)),\dots , (x_{n}, P(x_{n})) \}\).
\end{definition}


\begin{theorem}[Uniqueness of an interpolating polynomial]
  For any set \( \{ (x_0,P(x_0)),\dots , (x_{n}, P(x_{n})) \}\) of $n+1$
  point-value pairs such that all the $x_k$ values are distinct,
  there is a unique (single) polynomial $P(x)$ of degree-bound $n$ such that
  $y_k = P(x_k)$ for $k\in[n]$ 
\end{theorem}

\begin{proof}
  \sloppy{
  So let there be a polynomial $p(x) \neq 0$ of degree 
  $n$ and a set $ \{ (x_0,P(x_0)),\dots , (x_{n}, P(x_{n})) \}$
  of points representing it. Let us assume in contradication that there exists $q(x)\neq p(x)$ such that $p(x_0)=q(x_0), \dots, p(x_n)=q(x_n)$.
  That is, there exists another polynomial that is evaluated to the same points,
  making the representation non-unique.

  If such $q(x)$ exists, we can create a new polynomial $(p-q)(x)$. Note that this polynomial may be of degree up to $n$.
  $x_i$ is a root $0\le i\le n$ of $(p-q)(x)$, since \((p-q)(x_i)=p(x_i)-q(x_i)\), thus it has $n+1$ roots.
  This is a contradication to the fundamental theorem of algebra, that is, a polynomial $p(x)\neq 0$ 
  of degree at most $n$ has at most $n$ roots.}
\end{proof}

Now that we know that this representation is unique, I'll claim without proof that
there is a mapping that is both injective and surjective (bijective) from both representations.
Thus these both know representations are isomorphic.

\section{Polynomial Identity Testing (SZ)}\label{schawtz-zippel}
In this section, we talk about identity testing, 
and some issues involving this subject.

Given two polynomials $P(x)$ and $Q(x)$ of the same degree $d$, 
how do we prove these are the same polynomials? We can compare their coefficients. 
What if we don't have access to their coefficients, instead we have an oracle that outputs for any input integer $x_i$ the values of $P(x_i)$ and $Q(x_i)$?
We can calculate it and check $d$ points on each polynomial 
since each polynomial can be solely defined by its point-value representation\ref{polynomials:point-val-representation}.


Schwartz-Zippel \cite{SZ} made this notion much more efficient.
By choosing a random point for evaluating each polynomial, we can 
be almost entirely certain whether the polynomials are equal or not.

\begin{theorem}
 Let $P(x)$ be a non-zero polynomial of total degree $d$ over 
 an integral domain $R$. Let $S$ be a finite subset of $R$, and let $r$
 be a selected random number from $S$. Then
  $\Pr [P(x)=0]\le \frac{deg(p(x))}{|S|}$.
\end{theorem}
\begin{proof}
  $P(x)$ is a polynomial, from the fundamental theorem of algebra\cite{Aigner2010} it has 
 at most $d$ roots. Thus, choosing some random point $r$,
  $\Pr [P(r)=0]\le \frac{d}{|S|}$. 
\end{proof}


Given this notion, one can easily extrapolate this idea and create a method 
to check whether $P(x)\cdot Q(x)$ is equal to some given $Z(x)$: 
choose a random number $r$, and evaluate $P(r)\cdot Q(r) - Z(r)$.
if this expression equates to $0$, then $P(x)\cdot Q(x)=Z(x)$ with overwhelming 
probability (depending on the size of $Z_q$).

\subsection{\texorpdfstring{Troubled waters in $Z_q[X]/F(X)$}{Troubled waters in Zq[X]/F(X)}}
Given the above method, we can investigate new and upcoming ideas on numerous
problems (see  \cite{sz-usage2} and \red{fpir- to be published}).
The unlucky ones might encounter a troubling problem.
While SZ works quite well over fields and rings,
One would quickly find out that more complex algebraic structures do not 
bend so easily to this powerful technique.

For instance, consider the following quotient polynomial rings $R_q=Z_q[X]/(X^n+1)$,
choose two random polynomials from $R_q$, and try the method of testing above.
It is highly likely that for some random $r$, $P(r)\cdot Q(r) \ne (PQ \mod (X^n+1))(r)$.

This issue arises because not every point $r\in Z_q$ can be used to construct a
polynomial evaluation map $f_r$ (refer to \autoref{def:poly-eval}) 
over $R_q$ such that $f_r$ is a ring homomorphism. 
Consequently, certain values of $r$ will not be compatible with the Schwartz-Zippel method.
In other words, $f_r(P\cdot Q) \neq f_r(P)\cdot f_r(Q)$.


\paragraph{Finding valuable points}

Fortunately, there are some points $r\in \Z_q$ which can be used to create polynomial evaluation maps that 
are ring homomorphisms.
\begin{theorem}
  Let $R_q=Z_q[X]/F(X)$ and let $r\in Z_q$. If $F(r)=0$ then $f_r$ is a ring homomorphism.
\end{theorem}
\begin{proof}
  $f_r$ preserve the multiplicativeidentity: $f_r(1_{R_q}) = 1(r)=1$.
  Every $f_r$ preserver addition.
  preserving multiplication:
  Let there be $P(X), Q(X) \in R_q$. 
  By definition $P(X)\cdot Q(X) \mod F(X) = P(X)\cdot Q(X) \lfloor \frac{P(X)\cdot Q(X)}{F(X)} \rfloor \cdot F(X)$,
  so $f_r(P(X)\cdot Q(X))= f_r(P(X))cdot f_r(Q(X)) - f_r(\lfloor  \frac{P(X)\cdot Q(X)}{F(X)} \rfloor)\cdot f_r(F(X))$.
  Because $f_r$ is an evaluation map, $f_r(\lfloor  \frac{P(X)\cdot Q(X)}{F(X)} \rfloor)=z$ where is some unknown integral from $Z_q$.
  then $f_r(P(X)\cdot Q(X)) =P(r)\cdot Q(r) - z \cdot 0= P(r)\cdot Q(r)$, as wanted.
\end{proof}

So, by finding the roots of $F(X)$, we can identify potential values where 
the SZ algorithm continues to function effectively. 
However, this set is quite limited (consider the maximum number of roots of the polynomial $F(X)$).

\section{Polynomial addition}\label{polynomials:addition}
\section{Fast Polynomial Multiplications}\label{polynomial:multiplication}\label{section:fft}

We know how to add two polynomials as efficient as possible~\ref{polynomials:addition},
but the simple definition of polynomial multiplication isn't as efficient, and demand $O(n^2)$.
Many complex algorithms make use of polynomial multiplication, and as a result, the algortihm 
we are going to describe was announced as one of the top 10 Algorithms of 20th
Century by the IEEE magazine Computing in Science and Engineering.\cite{814652}


We begin by remembering the point-value representation~\ref{polynomials:point-val-representation}.
In this representation, once can perform both addition and multiplication between two polynomials 
by performing point to point operations (similar to addition in coefficient representation)\todo{provide proof?}.
As a result multiplication and addition in this representation cost $O(n)$.


The naive algorithm of moving from coefficient representation to point-value representation
requires evaluating the polynomials $n+1$ times, and as a result runs in $O(n^2)$.
Thus, to perform polynomial multiplication we'll begin with finding a clever and fast way to move between representations.
\subsection{FFT}
Luckily there is a way to move to and from both representations quickly using the \emph{roots of unity}.

\begin{definition}
An $n$th root of unity is a complex number $\omega$ such that
$\omega^n =1$.
\end{definition}

There are exactly $n$ roots of unity $e^{2\pi i k/n}$ for $k\in[n-1]$ 
($e^{iu}= \cos{u}+i \sin{u}$).

The value $\omega_n=e^{2\pi i/n}$ is the \emph{principal nth root of unity}.
all other roots of unity are powers of $\omega_n$:
 $\omega_n^0, \omega_n^1, \omega_n^2,\dots, \omega_n^{n-1}$.

FFT takes special advantage that these values have special properties.
\begin{lemma}[roots of unity]\label{roots-of-unity-property}
If $z$ is an nth root of unity, then $ a \equiv b \mod n \Rightarrow z^a = z^b$.
\end{lemma}
meaning that $\omega_n^a = \omega_n^b$. one can label specific points on the unit circle over $\C$.

\paragraph{Why is that important?}

In the simplest of cases, we are going to evaluate the polynomial over the following
points $ \{\omega_n^0, \omega_n^1, \omega_n^2,\dots, \omega_n^{n-1}\}$.
Let us view the polynomial evaluation over $n$ such points as a matrix, this presentation
might help the reader see some similarities:


$$
\begin{pmatrix}
  y_0\\
  y_1\\
  \vdots \\
  y_{n-1}\\
\end{pmatrix}
=
\begin{pmatrix}
  1 & x_0 & x_0^2 ,\dots\\
  1 & x_1 & x_1^2, \dots,\\
  \vdots & \vdots & \vdots & \ddots & \\
  1 & x_{n-1} & x_{n-1}^2, \dots,\\
 \end{pmatrix}
 \cdot 
\begin{pmatrix}
  a_0\\
  a_1\\
  \vdots \\
  a_{n-1}\\
\end{pmatrix}
 $$

 Now, let us assign the values:

 \label{dft-matrix}
 $$
\begin{pmatrix}
  p(\omega_n^0)\\
  p(\omega_n^1)\\
  \vdots \\
  \omega_n^2\\
\end{pmatrix}
=
\begin{pmatrix}
  1 & (\omega_n^0)^1 & (\omega_n^0)^2 ,\dots ,(\omega_n^0)^{n-1}\\
  1 & (\omega_n^1)^1 & (\omega_n^1)^2, \dots, (\omega_n^1)^{n-1}\\
  \vdots & \vdots & \vdots & \ddots & \\
  1 & (\omega_n^{n-1}) & (\omega_n^{n-1})^2, \dots,(\omega_n^{n-1})^{n-1}\\
 \end{pmatrix}
 \cdot 
\begin{pmatrix}
  a_0\\
  a_1\\
  \vdots \\
  a_{n-1}\\
\end{pmatrix}
 $$

 The keen looker can see that we can reuse values in this matrix.
 for example, let us look at $(\omega_n^{n-1})^{n-1}$ using \ref{roots-of-unity-property}
 $(\omega_n^{n-1})^{n-1} = (\omega_n)^{n*n-2n+1}=(\omega_n)^{0+0+1}$.  

 Thus we gain a matrix with heavy symmetry, which the FFT can use.

\paragraph{fft main idea}
The FFT method employs a divide-and-conquer strategy, 
using the even-indexed and odd-indexed coefficients of $A(x)$
separately to define the two new polynomials $A^{[0]}(x)$ and $A^{[1]}(x)$ of degree-bound $n/2$.

\begin{align}
  &A^{[0]}(x) =  a_0 +a_2x + a_4x^2 + \dots + a_{n-2}x^{n/2-1}\\
  &A^{[1]}(x) =  a_1 +a_3x + a_5x^2 + \dots + a_{n-1}x^{n/2-1}.
\end{align}


It follows that $A(x)=A^{[0]}(x^2) + x\cdot A^{[1]}(x^2)$.
\begin{examplebox}{Example: splitting a degree-8 polynomial}
Take $A(x) = 1 + 2x + 3x^2 + 4x^3 + 5x^4 + 6x^5 + 7x^6 + 8x^7$ with $n = 8$.
The even-indexed coefficients are $a_0=1,\, a_2=3,\, a_4=5,\, a_6=7$, and the odd-indexed coefficients are $a_1=2,\, a_3=4,\, a_5=6,\, a_7=8$, giving
$$A^{[0]}(x) = 1 + 3x + 5x^2 + 7x^3, \qquad A^{[1]}(x) = 2 + 4x + 6x^2 + 8x^3.$$
We can verify the identity $A(x) = A^{[0]}(x^2) + x \cdot A^{[1]}(x^2)$:
\begin{align*}
  A^{[0]}(x^2) + x \cdot A^{[1]}(x^2)
  &= (1 + 3x^2 + 5x^4 + 7x^6) + x(2 + 4x^2 + 6x^4 + 8x^6)\\
  &= 1 + 2x + 3x^2 + 4x^3 + 5x^4 + 6x^5 + 7x^6 + 8x^7 = A(x). \checkmark
\end{align*}
The two subpolynomials each have degree bound $n/2 = 4$, illustrating how the problem halves at each recursive step.
\end{examplebox}

Thus, the problem of evaluating $A(x)$ at 
$\omega_n^0, \omega_n^1, \omega_n^2,\dots, \omega_n^{n-1}$ 
reduces to evaluating the degree-bound $n/2$ polynomials  $A^{[0]}(x)$ and $A^{[1]}(x)$ at the points
$$(\omega_n^0)^2, (\omega_n^1)^2,\dots, (\omega_n^{n-1})^2$$
and then combining the results according to $A(x)=A^{[0]}(x^2) + x\cdot A^{[1]}(x^2)$.


Using the properties of the root of unity, 
the above values consist not of $n$ distinct values but only of the $n/2$
 complex ($n/2$)th roots of unity, with each root occurring exactly twice!
These subproblems have the same form as the original problem but are half the size.
We can continue in this manner until we evaluate the full polynomial on $n$ distinct values.

\begin{examplebox}{Example: DFT matrix with permuted coefficient vector ($n=4$)}
Reordering the coefficient vector as $[a_0, a_2, a_1, a_3]^T$ (even indices first, then odd), the DFT evaluation becomes
$$
\begin{pmatrix} A(\omega_4^0) \\ A(\omega_4^1) \\ A(\omega_4^2) \\ A(\omega_4^3) \end{pmatrix}
=
\begin{pmatrix}
  (\omega_4^0)^0 & (\omega_4^0)^2 & (\omega_4^0)^1 & (\omega_4^0)^3 \\
  (\omega_4^1)^0 & (\omega_4^1)^2 & (\omega_4^1)^1 & (\omega_4^1)^3 \\
  (\omega_4^2)^0 & (\omega_4^2)^2 & (\omega_4^2)^1 & (\omega_4^2)^3 \\
  (\omega_4^3)^0 & (\omega_4^3)^2 & (\omega_4^3)^1 & (\omega_4^3)^3
\end{pmatrix}
\begin{pmatrix} a_0 \\ a_2 \\ a_1 \\ a_3 \end{pmatrix}
=
\begin{pmatrix}
  \omega_4^0 & \omega_4^0 & \omega_4^0 & \omega_4^0 \\
  \omega_4^0 & \omega_4^2 & \omega_4^1 & \omega_4^3 \\
  \omega_4^0 & \omega_4^4 & \omega_4^2 & \omega_4^6 \\
  \omega_4^0 & \omega_4^6 & \omega_4^3 & \omega_4^9
\end{pmatrix}
\begin{pmatrix} a_0 \\ a_2 \\ a_1 \\ a_3 \end{pmatrix}.
$$
Row $j$ evaluates $A$ at $\omega_4^j$, so each entry is $(\omega_4^j)^m$ where $m$ is the exponent of $x$ for that coefficient.
The first row ($j=0$) collapses to all $\omega_4^0$ since $(\omega_4^0)^m = 1$ for any $m$.
The left half of each row (columns for $a_0, a_2$) contains only even powers of $\omega_4$,
and the right half (columns for $a_1, a_3$) contains only odd powers.
This separation is exactly what allows the two size-$n/2$ DFTs to be computed independently.
\end{examplebox}

\subsection{putting it all together}
As we've seen we can create a matrix with heavy symmetry, called the DFT matrix. 
now, we'd want to capitalize on that and create a recursive algorithm.
To do so, instead of the regular polynomial assignment we've seen above using a matrix,  
we'll shuffle things a bit according to the divide-and-conquer approach from above

\begin{examplebox}{Example: factoring the $n=4$ DFT matrix}
Starting from the permuted matrix in the previous example, notice that the left two columns
(even coefficients $a_0, a_2$) repeat identically in the top and bottom halves, while the right two columns
(odd coefficients $a_1, a_3$) appear with opposite sign between the halves.
Grouping these blocks, the matrix factors as
$$
\left(\begin{array}{cc|cc}
  \omega_4^0 & \omega_4^0 & \omega_4^0 & \omega_4^0 \\
  \omega_4^0 & \omega_4^2 & \omega_4^1 & \omega_4^3 \\
  \hline
  \omega_4^0 & \omega_4^4 & \omega_4^2 & \omega_4^6 \\
  \omega_4^0 & \omega_4^6 & \omega_4^3 & \omega_4^9
\end{array}\right)
=
\begin{pmatrix} I_2 & D_2 \\ I_2 & -D_2 \end{pmatrix}
\begin{pmatrix} F_2 & 0 \\ 0 & F_2 \end{pmatrix}
$$
where $D_2 = \begin{pmatrix} \omega_4^0 & 0 \\ 0 & \omega_4^1 \end{pmatrix}$ are the twiddle factors, and
$F_2 = \begin{pmatrix} \omega_2^0 & \omega_2^0 \\ \omega_2^0 & \omega_2^1 \end{pmatrix}$
is the size-2 DFT matrix applied independently to the even and odd coefficients.
Multiplying out confirms the factorization:
$$
\begin{pmatrix} I_2 & D_2 \\ I_2 & -D_2 \end{pmatrix}
\begin{pmatrix} F_2 & 0 \\ 0 & F_2 \end{pmatrix}
=
\begin{pmatrix} F_2 & D_2 F_2 \\ F_2 & -D_2 F_2 \end{pmatrix}.
$$
This is exactly the recursive structure: compute $F_2$ on the even and odd halves separately, then combine with the twiddle factors in $D_2$.
\end{examplebox}


$$
F_n \cdot p(x) =
\begin{pmatrix}
  I_{ \frac{n}{2}} & D_{\frac{n}{2}} \\
  I_{ \frac{n}{2}} & -D_{\frac{n}{2}}
 \end{pmatrix}
 \cdot 
 \begin{pmatrix}
  F_{ \frac{n}{2}} & 0 \\
  0 & F_{ \frac{n}{2}} 
\end{pmatrix}
\cdot 
\begin{pmatrix}
  p_{\text{even coefficient}} \\
  p_{\text{odd coefficient}}
\end{pmatrix}
 $$

 where $F_n$ is the matrix of assignment we've seen in ~\ref{dft-matrix}, and 
 $D_n$ is the following diagonal matrix $$
 \begin{pmatrix}
  1 & 0 &0  & \dots& 0 \\
  0 & \omega_n^1 &0  & \dots& 0 \\
  1 & 0  &\omega_n^2  & \dots& 0 \\
  \vdots & \vdots & \vdots & \vdots &  \\
  0 & 0  &0  & \dots& \omega_n^n-1 \\
 \end{pmatrix} $$


 We can continue this process recursively, to get what we want.
 In each recursive level, the algorithm does at most $O(n)$ work (i.e., multiplying by sparse matrices),
  and due to the logarithmic depth of the recursion, we get $O(n\log{n})$.
x
\label{algorithm:fft}
  \begin{lstlisting}[language=Python,
    caption={fft algorithm in python}
  ]
import numpy as np

def fft(p: np.ndarray) -> np.ndarray:
    n = len(p)
    if n == 1:
        return p
    omega = np.exp(2 * np.pi * 1j / n)

    p_even, p_odd = p[::2], p[1::2] 
    ye, yo = fft(p_even), fft(p_odd)

    # mult the structured left matrix with the recursive result.
    # O(n) work comes from here.
    y = np.zeros(n, dtype=np.complex128)
    for k in range(n // 2):
        y[k] = ye[k] + omega ** k * yo[k]
        y[k + n // 2] = ye[k] - omega ** k * yo[k]
    return y
    \end{lstlisting}
    
    \begin{remark}[Summary]
      Notice that to achieve this result, we needed to reduce the polynomial into
      even and odd parts, and then adjust the assignment matrix using
      the properties of roots of unity. Each part is important in its own way.
    \end{remark}

% \begin{figure}[ht]
% \centering
% \begin{tikzpicture}[>=stealth, scale=0.95]
%   \def\dy{0.65}  % vertical spacing between rows
%   \def\st{2.3}   % horizontal spacing between stages

%   % Input nodes and labels (bit-reversed order for DIT-FFT)
%   \foreach \i/\lbl in {0/{a_0},1/{a_4},2/{a_2},3/{a_6},4/{a_1},5/{a_5},6/{a_3},7/{a_7}} {
%     \node[left, font=\scriptsize] at (0, {-\i*\dy}) {$\lbl$};
%     \filldraw (0, {-\i*\dy}) circle (1.5pt);
%   }
%   % Intermediate nodes after stages 1 and 2
%   \foreach \col in {1,2} {
%     \foreach \i in {0,...,7}
%       \filldraw ({\col*\st}, {-\i*\dy}) circle (1.5pt);
%   }
%   % Output nodes and labels
%   \foreach \i in {0,...,7} {
%     \filldraw ({3*\st}, {-\i*\dy}) circle (1.5pt);
%     \node[right, font=\scriptsize] at ({3*\st}, {-\i*\dy}) {$A(\omega^{\i})$};
%   }

%   % ---- Stage 1: butterflies (0,1),(2,3),(4,5),(6,7) ----
%   \foreach \ra/\rb in {0/1, 2/3, 4/5, 6/7} {
%     \draw[gray!40]   (0,{-\ra*\dy}) -- (\st,{-\ra*\dy});
%     \draw[gray!40]   (0,{-\rb*\dy}) -- (\st,{-\rb*\dy});
%     \draw[blue!65]   (0,{-\ra*\dy}) -- (\st,{-\rb*\dy});
%     \draw[blue!65]   (0,{-\rb*\dy}) -- (\st,{-\ra*\dy});
%   }

%   % ---- Stage 2: butterflies (0,2),(1,3),(4,6),(5,7) ----
%   \foreach \ra/\rb in {0/2, 1/3, 4/6, 5/7} {
%     \draw[gray!40]   (\st,{-\ra*\dy}) -- ({2*\st},{-\ra*\dy});
%     \draw[gray!40]   (\st,{-\rb*\dy}) -- ({2*\st},{-\rb*\dy});
%     \draw[blue!65]   (\st,{-\ra*\dy}) -- ({2*\st},{-\rb*\dy});
%     \draw[blue!65]   (\st,{-\rb*\dy}) -- ({2*\st},{-\ra*\dy});
%   }

%   % ---- Stage 3: butterflies (0,4),(1,5),(2,6),(3,7) ----
%   \foreach \ra/\rb in {0/4, 1/5, 2/6, 3/7} {
%     \draw[gray!40]   ({2*\st},{-\ra*\dy}) -- ({3*\st},{-\ra*\dy});
%     \draw[gray!40]   ({2*\st},{-\rb*\dy}) -- ({3*\st},{-\rb*\dy});
%     \draw[blue!65]   ({2*\st},{-\ra*\dy}) -- ({3*\st},{-\rb*\dy});
%     \draw[blue!65]   ({2*\st},{-\rb*\dy}) -- ({3*\st},{-\ra*\dy});
%   }

%   % ---- Twiddle factor labels ----
%   % Stage 1: all ω^0
%   \node[above, font=\scriptsize, blue!80] at ({0.5*\st}, 0.15) {$\omega^0$ (all)};

%   % Stage 2
%   \node[font=\scriptsize, blue!80] at ({1.5*\st}, {-1*\dy})   {$\omega^0$};
%   \node[font=\scriptsize, blue!80] at ({1.5*\st}, {-2*\dy})   {$\omega^2$};
%   \node[font=\scriptsize, blue!80] at ({1.5*\st}, {-5*\dy})   {$\omega^0$};
%   \node[font=\scriptsize, blue!80] at ({1.5*\st}, {-6*\dy})   {$\omega^2$};

%   % Stage 3
%   \node[font=\scriptsize, blue!80, right=2pt] at ({2.5*\st}, {-2*\dy}) {$\omega^0$};
%   \node[font=\scriptsize, blue!80, right=2pt] at ({2.5*\st}, {-3*\dy}) {$\omega^1$};
%   \node[font=\scriptsize, blue!80, right=2pt] at ({2.5*\st}, {-4*\dy}) {$\omega^2$};
%   \node[font=\scriptsize, blue!80, right=2pt] at ({2.5*\st}, {-5*\dy}) {$\omega^3$};

%   % ---- Stage labels ----
%   \node[above, font=\small, anchor=south] at ({0.5*\st}, 0.45) {Stage 1};
%   \node[above, font=\small, anchor=south] at ({1.5*\st}, 0.45) {Stage 2};
%   \node[above, font=\small, anchor=south] at ({2.5*\st}, 0.45) {Stage 3};

%   % Vertical dividers
%   \foreach \col in {1,2,3}
%     \draw[densely dashed, gray!40] ({\col*\st}, 0.4) -- ({\col*\st}, {-7.4*\dy});

%   \node[above, font=\tiny, gray, anchor=south] at (0, 0.4) {bit-rev.};
% \end{tikzpicture}
% \caption{FFT butterfly diagram for $n=8$ (Cooley--Tukey, decimation-in-time).
% Inputs are in bit-reversed order (left); three stages of butterfly operations produce
% the DFT values $A(\omega^0),\dots,A(\omega^7)$ (right), where $\omega = e^{2\pi i/8}$.
% Each pair of blue crossing lines is one butterfly: the twiddle factor $\omega^k$ labels
% the crossing midpoint and multiplies the lower input before the $\pm$ combination.
% The stride doubles at each stage: Stage~1 connects adjacent rows, Stage~2 skips one,
% and Stage~3 spans four rows.}
% \end{figure}

\subsection{Inverse FFT}\label{ifft}

The DFT matrix is invertible.
Thus one can solve for the vector of coefficients,
by simply inverting the matrix in~\ref{dft-matrix}.

Fortunately, with little tricks we can adjust the inverse of the DFT matrix such that
we can use the same algorithm as above, but instead of
$\omega^k$ we'll use $\omega^{-k}$, and after we are done with the FFT
algorithm, we'll need to normalize it by dividing every coefficient element
by $N$ the size of the polynomial.

 \begin{corollary}
  For FFT, DFT, and NTT we must use specific values to gain speedy interpolation and
  evaluation of the polynomial. 
 \end{corollary}{Conclusin:}
 
\subsection{Number Theoretic Transform (NTT)}\label{ntt}

NTT is a specialized version of the discrete Fourier transform,
in which the coefficients of the polynomial are members of a finite field (or ring)
 containing the right roots of unity.
It can be viewed as an exact version of the complex DFT, avoiding 
round-off errors for exact convolutions of integer sequences.
While Gauss used similar techniques already in~\cite{gaussfft}, laying the groundwork 
for modern FFT algorithms to compute the DFT, and therefore the NTT, 
is usually attributed to Cooley and Tukey's seminal paper \cite{fft}.

NTT can use the same FFT algorithms but requires a 
different DFT matrix~\ref{dft-matrix}.

This fact lies in the foundational theorem of NTT:
\begin{theorem}[\cite{ntt}]
  A length $N$ transform having the DFT structure will
  implement cyclic convolution if and only if there exists an 
  inverse of N and an element $\alpha$,
  a root of unity of order $N$ , i.e., $N$ is the least 
  positive integer such that $\alpha^N =1$.
\end{theorem}
This is a very general result applying to both \emph{rings and fields} that are
finite or infinite.

It is important to note that $N$ also applies to the length of our polynomial.

\subsubsection{Preliminaries}
I will describe everything in terms of fields, but essentially all of this can be easily generalised to rings 
using the Chinese Remainder Theorem (CRT), as shown below.
\paragraph{ Finding Roots of Unity (In Field)}
As described above, an $n$-th root of unity, is an element $\alpha\in Z_q$ 
such that $\alpha^n =1  \mod q $, and for every $0<k<n$ $\alpha^k \neq 1 \mod q$.

\paragraph{Finding roots of unity in a field}

\begin{theorem}
  Let $g\in GF(q)$ be a generator, thus
  $\omega_0=g^{(q-1)/n}$ is an $n$-th root of unity.
\end{theorem}
\begin{proof}
  \sloppy{
  relying on \autoref{fermats-little-theorem}, we know that $g^{q-1}=1 \mod q$.
  So $\omega_0^n=(g^{(q-1)/n})^n=1$, now we just need to explain away why $\omega_0^k \ne 1$ for 
  every $0<k<n$.
  Let us assume in contradiction that there exists some $k<n$ such that $\omega_0^k=1$, then 
  $\omega_0^k=g^{(q-1/n)\cdot k}=1$, more specifically, $g^c=1$ for $c<q-1$. thus, we have a 
  cycle $(1,g,g^1,...g^{c-1})$, which is less than $q-1$ elements, meaning that $g$ is not 
  a generator, in contradiction.
  }
\end{proof}

Now that we found the first root of unity,
we can find the rest by computing the powers of $\omega_0$ up to $n$.


But how do we find the generator to begin with? 
There is a randomised algorithm to find a generator in an expected 
$\approx O(\sqrt{q-1})+O(\log^2{q}\cdot \log (\log(q-1)))$ time for small primes.
For larger primes it'll longer, but it is still feasible ($O(e^{\log q-1})$).

The algorithm itself is based on factorising $q-1$ into prime factors, and then
check that a random number $g$ is a generator by ensuring $g^{(p-1)/p_i} \not\equiv 1$
for each $p_i$ prime factor of $q-1$ (where $q$ is the field's prime order).


\paragraph{Neat Trick}
\todo{This section would have saved me days of work, had I known it before.}
One can make use of these roots of unity to find the roots of $X^n+1$ and $X^n-1$.
Finding the roots of $X^n-1$ is from the definition! 
Every n-th root of unity is a root of $X^n-1$: $(\omega^k)^n=(\omega^n)^k=1^k=1$.
Substitute $X$ with any n-th root of unity, and you'll get $1-1=0$.

Finding the roots of $X^n+1$ is a bit more tricky,
but it can be done by finding the roots of $X^{2n}-1$ (which should be easy
as we've shown above):
First let's recall that $X^{2n}-1=(X^n+1)(X^n-1)$, then we notice that
we have $n$ roots of $X^n-1$,
and $n$ roots of $X^n+1$. So every one of the roots of $X^{2n}-1$ that is
not a root of $X^n-1$ is a root of $X^n+1$.
How do we find the roots of $X^{2n}-1$? Those are the $2n$-th roots of unity, 
which are easy to find!

\paragraph{Chinese Remainder Theorem (CRT) For Dummies}
CRT is a way of representing elements $x\in Z_q$ (
  where $q=p_1,p_2,\dots,p_k$ and $p_i\in\text{Primes}$)
such that the $x_{CRT}$ is a single and unique representation of $x$ using the 
remainders of each of the prime factors constructing $q$:

$$ x_{CRT} = (x\mod p_1, x\mod p_2 ,\dots x\mod p_k ) $$



Because it is a unique representation, given $x_{CRT}$ and the list of
 primes $(p_1,p_2,\dots,p_k)$ there is an algorithm to solve for $x \in Z_q$
 (i.e., find $x \mod q$).


Another important thing to notice is that if  $x=1 \mod q$
 then $x_{CRT}= (1,1,1,\dots,1)$.


There is more to it honestly, but for our cause, it's enough.

\subsubsection{Putting it all together}

\paragraph[NTT Over Finite Fields]{ NTT Over Finite Fields} \label{ntt-field}
When working over a field $GF(p)$ such that $p$ is prime,
the maximal $N$, the array of coefficients representing a
polynomial, is $N_{max}=p-1$, and $N$ is a valid size for NTT
 in this field iff $N|p-1$ (\cite[III. number theoretic transforms]{ntt}). 


To compute the NTT over $GF(p)$, find the $n$th root of unity in $GF(p)$,
 then run any FFT algorithm mod $p$ using the roots
of unity for each level of the recursion.

That is, find $\omega_n=\alpha$ such that $\alpha^N=1 \mod p$, 
then run the (almost) same algorithm in \nameref{algorithm:fft}, 
but change line 7 such that omega is now the $n$-th root of unity in $GF(p)$, and
all operations in the algorithm are now $\mod p$.

\paragraph{iNTT}
Just like in \nameref{ifft}, but to calculate $\omega^-1$, we'll
use the algorithm shown in \autoref{sec:finite-field:extended-euclidean} 
to find the inverse of every element ($x^-1$).



\paragraph[short]{NTT over Rings}
Let $Z_M$ represent a ring of integers ${0,1,2,\dots , M-1}$,
with arithmetics mod $M$.
Where $M$ is constructed the following unique prime factorization
$M=p_{1}^{r_1}\cdot  p_{2}^{r_2} \cdots p_{\ell}^{r_\ell}$
and $p_i$'s are distinct primes (see \cite[NTT sec. III][]{ntt}
 for further explanations).

For NTT to work, the following must hold $N|L(M)$
where $L(M)$ is the greatest common
divisor (gcd) of the $(p_{i}-1)$. That is, 
$L(M):= gcd\{p_1-1,p_2-1,\dots\}$

essentially it means we want $ N$ to divide $p_i-1$.

\begin{corollary}
  $L(M)=N_{max}$. i.e., it defines the maximal $N$ 
  such that we can apply NTT.  
\end{corollary}

Now, Once $M$ and $N$ are known, to perform NTT we just need to be able to find 
$N$-th root of unity. but it's quite hard to find one in $Z_M$.
Luckily, we know it is feasible to do so in a prime field. 
So, let's find $\alpha_i$, the $N$-th root of unity for every $p_i$.
Once we find all of these roots of unity, we know that $\alpha_i^N=1 \mod p_i$ for every $i$.
In other words, $N$ is the least positive integer such that
\begin{align*}
  &\alpha_{CRT}^N = (\alpha_1^N \mod p_1=1, \alpha_2^N \mod p_2=1,\dots)= (1,1,\dots)
\end{align*}

and according to CRT $\alpha^N=1 \mod M$. Thus, solving for $\alpha$ 
means we'll be able to find the $N$-th root of unity in $Z_M$.

The algorithm remains the same as in \nameref{ntt-field} from here on.

\paragraph{Special Features}
Applying the NTT transform provides a cyclic convolution, computing 
$c = a \cdot b \mod (X^n + 1)$  with two polynomials $a$ and $b$ would require 
applying the NTT of length $2n$ and thus $n$ zeros to be appended to each input;
this effectively doubles the length of the inputs and also requires the computation of 
an explicit reduction modulo $X^n +1$. 
To avoid these issues, one can exploit the \emph{negative wrapped convolution} \cite{negantt}.

This exploitation allows us to perform $NTT$ without expanding to $2n$ parameters!
meaning we can send less bandwidth over the network, and perform fewer computations too. 
This trick can be seen in \cite{SEAL}.

\section{Fast Polynomial Division via Reversal and Newton Iteration}
\begin{insight}
The algorithm for fast polynomial division attempts to find the inverse of the divisor polynomial. To do so, it modifies it such that one can apply the Newton iteration.
Using the Newton iteration with specific loop invariant allows us to find the inverse of the modified divisor polynomial. 
Once done, one can modify the result such that it conforms to the quotient polynomial we are looking for.

Algorithm: reverse coefficients, apply Newton iteration to find the inverse of the reversed divisor, multiply this with the reversed dividend, and reverse the result to get the quotient.
\end{insight}

\red{Notice that this section is a somewhat explanation for section 9.1 of Modern Computer Algebra}
Let $A(x), B(x) \in \mathbb{F}[x]$ with $\deg(A)=n$, $\deg(B)=m$, and assume
the leading coefficient of $B(x)$ is invertible in $\mathbb{F}$. We want to compute
the quotient $Q(x)$ and remainder $R(x)$ such that
\[
A(x) = Q(x) B(x) + R(x), \quad \deg(R) < m.
\]

\paragraph{Reversal.}
Define the \emph{reversal} of a polynomial $f(x)=f_0 + f_1 x + \dots + f_d x^d$ of degree $d$
by
\[
\operatorname{rev}_L(f)(x) \;=\; x^{L-1} f\!\left(x^{-1}\right) \bmod x^L,
\]
which essentially reverses the coefficients up to length $L$.

\begin{examplebox}
  Let $f(x) = 1 + 2x + 3x^2 + 4x^3 + 5x^4$, so $\deg(f)=4$.
  First substitute $x^{-1}$:
  \[ f(x^{-1}) = 1 + 2x^{-1} + 3x^{-2} + 4x^{-3} + 5x^{-4}. \]
  Then multiply by $x^{L-1}$:
  \[ x^{L-1}f(x^{-1}) = x^{L-1} + 2x^{L-2} + 3x^{L-3} + 4x^{L-4} + 5x^{L-5}, \]
  and the $\bmod\, x^L$ step keeps only terms of degree $0$ through $L-1$.
  \begin{itemize}
    \item $L=5$ (equals $\deg(f)+1$): all terms have degree $0,\dots,4$, nothing is dropped.
      \[ \operatorname{rev}_5(f) = 5 + 4x + 3x^2 + 2x^3 + x^4. \]
      Coefficients $(1,2,3,4,5)$ are fully reversed to $(5,4,3,2,1)$.
    \item $L=7$ (greater than $\deg(f)+1$): terms have degree $2,\dots,6$, all below $L$, so nothing is dropped but the result is zero-padded at the low end.
      \[ \operatorname{rev}_7(f) = 5x^2 + 4x^3 + 3x^4 + 2x^5 + x^6. \]
    \item $L=3$ (less than $\deg(f)+1$): terms with degree $< 0$ (from $4x^{L-4}=4x^{-1}$ and $5x^{L-5}=5x^{-2}$) are discarded. Only the first $L$ coefficients of $f$ survive, reversed.
      \[ \operatorname{rev}_3(f) = 3 + 2x + x^2. \]
      The leading coefficients $4$ and $5$ are lost.
  \end{itemize}
\end{examplebox}

We want to find $Q(x)$, and we know that $Q(x) = A(x)\cdot B(x)^{-1}$, so we can find $Q$ by finding the inverse of $B$ and multiplying it by $A$.
$Q(x)$ has degree $k-1$ with $k=n-m+1$, so we can find $Q$ by finding the inverse of $B$.

What is the inverse of $B$? It is the polynomial $B^{-1}$ such that $B(x) \cdot B^{-1}(x) = 1$.
\begin{notice}
  How does reversing $B(X)$ help us find the inverse of $B$? It doesn't! However,
  it ensures that the leading coefficient of $B$ becomes the constant term of $\operatorname{rev}_{m+1}(B)$, which is crucial for the Newton iteration to work and  find the inverse of $\operatorname{B}_{m+1}(f)(x)$. 
\end{notice}

Given the reverse operator, we apply it to each of the polynomials, allowing us to find the reversed quotient $\operatorname{rev}_k(Q)$ (for $k=n-m+1$ the degree of $Q$) by multiplying $\operatorname{rev}_{n+1}(A)$ with the inverse of $\operatorname{rev}_{m+1}(B)$ (once we find it).
\[
\operatorname{rev}_k(Q)(x) \;=\; \operatorname{rev}_{n+1}(A)(x) \cdot
   \bigl( \operatorname{rev}_{m+1}(B)(x) \bigr)^{-1} \bmod x^k.
\]

% (Since $Q(x) = A(x)\cdot B(x)^{-1}$.)
% This reversal operation will later be used to \emph{convert the problem of dividing}
% high-degree polynomials into the problem of inverting a series with a nonzero
% constant term}, which can be solved efficiently by Newton iteration(see next). In other words,
% reversal ensures that the divisor's leading coefficient becomes the constant term
% of $\operatorname{rev}_{m+1}(B)$, so its inverse exists modulo $x^k$ and can be
% computed in $O(M(k))$ time (i.e., $M(k)$ describe the cost of multiplication. using NTT it is $k\cdot \log k$).

% In other words, the newtwon iteration, which we describe next demands the polynomial has a constant, invertible, nonzero constant. Otherwise, it doesn't converge (don't ask me why)!

\paragraph{Series inversion via Newton iteration.}
Newton iteration is a technique to find roots of polynomials (similar to gradient descent). 
Given some good starting point, each iteration approaches the root (if exists), up to some precision $k$.

In our case, we modify technique to find a polynomial $T(X)$ to serve as the inverse of our input (reversed $B(X)$ which we denote as $C(X)$ to avoid confusion).
That is, we build the function $F(T)=T(X)^{-1}-C(X)$ and seek its root (when $T(X)^{-1}=C(X)$).
This specific Newton iteration modification operates over the ring $\mathbb{F}/(x^k)$.
%  that is, if we find $T(X)^{-1}$ that can serve as a root, we know that $T(X)^{-1}=C(X)$, or in other words $T(X)$ is the inverse of $C(X)$. 
\begin{notice}
    You might be wondering why we choose $F(T)=T(X)^{-1}-C(X)$, and not something else. 
    For example, a more straightforward $F(T)$ would be $F(T)=1-T(x)C(X)$, but this leads to 
    an issue with the iteration. Try to apply this to the iteration after you understand how it works, and you'll get stuck.
\end{notice}

So we begin with $x_1= T_1(X)=c_0^{-1}$ (the inverse of the constant term of $C(X)$), 
then compute each step using the formula of the Newton iteration technique:
$T_{i+1}= T_i - \frac{F(T_i)}{F'(T_i)}= T_i - \frac{1/T_i(X)-C(x)}{-1/T_i(X)^2}$
So the computation in each step is $T_{i+1}= 2T_i(x)-C(X) \cdot T_i(X)^2$

Thus, Given $C(x)$ wih nonztero constant term, its inverse series modulo $x^k$ can be
computed by Newton iteration:
\[
T_{1}(x) = c_0^{-1}, \qquad
T_{2\ell}(x) = T_\ell(x) \,\bigl(2 - C(x)T_\ell(x)\bigr) \bmod x^{2\ell}.
\]
This converges doubly in the precision, so after $O(\log k)$ iterations we obtain
$C(x)^{-1} \bmod x^k$.
It is important to notice, this function uses polynomial multiplication, but the overall work remains in $O(M(n))$.

\begin{lstlisting}[language=Python,
  caption={Newton iteration for polynomial series inverse},
]

def series_inverse(C, k):
    """
    Input: polynomial $C(x)$ with C[0] != 0
           integer k = target precision
    Output: polynomial $T(x)$ such that $C(x)*T(x) = 1 mod x^k$
    """
    # initial inverse modulo x
    T = [ inverse(C[0]) ]  
    ell = 1

    while ell < k:
        m = min(2*ell, k)

        # U = C*T mod x^m
        U = multiply(C, T, m)

        # U = 2 - U mod x^m
        U[0] = 2 - U[0]
        for i in range(1, m):
            U[i] = -U[i]

        # T = T*U mod x^m
        T = multiply(T, U, m)

        ell = m

    return T
\end{lstlisting}

proof of correctness is via induction. (See Modern Computer Algebra book, section 9.1, theorm 9.2).
and runtime is around $4M(n)$, so if we use NTT, it should be aroud four times the cost of regular NTT multiplication.

\paragraph{Why Newton Iteration Applies.}
In polynomial division we know in advance the degree bound of the quotient:
if $\deg(A)=n$ and $\deg(B)=m$, then $\deg(Q)=n-m$. Hence we only require
the inverse of $\operatorname{rev}_{m+1}(B)(x)$ modulo $x^k$, where
$k = n-m+1$. This gives Newton iteration a natural stopping criterion:
compute the inverse series only up to precision $x^k$.



\paragraph{Division Algorithm.}
\begin{enumerate}
  \item Compute $A^*(x) = \operatorname{rev}_k(A)(x)$ and $B^*(x) = \operatorname{rev}_{m+1}(B)(x)$.
  \item Compute $T(x) = (B^*(x))^{-1} \bmod x^k$ using Newton iteration.
  \item Compute $Q^*(x) = A^*(x)\,T(x) \bmod x^k$.
  \item Recover the quotient by reversal: $Q(x) = \operatorname{rev}_k(Q^*)(x)$.
  \item Finally, compute the remainder as $R(x) = A(x) - Q(x)B(x)$.
\end{enumerate}

\paragraph{Complexity.}
If polynomial multiplication of degree-$n$ polynomials costs $M(n)$ operations,
then this division runs in $O(M(n))$ time. Using FFT/NTT-based multiplication,
this yields quasi-linear complexity in $n$.
\section{Half-GCD}\label{algorithm:half-GCD}
\red{this section was written with codex 5.3, needs to be reviewed further when I am not so tired.}
Half-GCD is a divide-and-conquer version of Euclid's algorithm.
The usual Euclidean algorithm removes one quotient at a time:
\[
(a,b) \mapsto (b, a - qb).
\]
Half-GCD keeps the same mathematics, but tries to process \emph{many} Euclid steps at once,
using only the high half of the inputs to predict those quotients.

\paragraph{Why this helps.}
When $a$ and $b$ are large (big integers, or high-degree polynomials),
the expensive part is repeatedly dividing large objects.
If we can reduce the size by about half in one recursive call,
we get a much faster algorithm asymptotically.
For polynomials over a field, this leads to roughly
\[
O(M(n)\log n)
\]
for GCD (and extended GCD), where $M(n)$ is multiplication cost.

\paragraph{Core viewpoint: Euclid as matrix updates.}
Each Euclid step with quotient $q$ can be written as
\[
\begin{pmatrix}a'\\ b'\end{pmatrix}
=
\begin{pmatrix}0 & 1\\ 1 & -q\end{pmatrix}
\begin{pmatrix}a\\ b\end{pmatrix}.
\]
So a block of Euclid steps is just multiplication by one $2\times 2$ matrix.
Half-GCD computes such a matrix from the \emph{most significant half}
of the input, then applies it to the full input.
This usually drops the size dramatically while preserving the same gcd.

\paragraph{Intuition for correctness (without proof).}
There are two simple facts:
\begin{itemize}
  \item Every Euclid step preserves the gcd.
  \item A product of Euclid-step matrices is still a valid gcd-preserving transform.
\end{itemize}
Half-GCD only groups these valid steps into larger chunks.
Even if a guessed chunk from the high half is imperfect,
we can safely correct with one ordinary Euclid step and continue.
So correctness comes from doing exactly the same legal transformations as Euclid,
just in larger batches.

\paragraph{How to think about one recursive level.}
Suppose inputs have size $n$ (bit-length for integers, degree for polynomials).
\begin{enumerate}
  \item Keep only the high $\approx n/2$ part of each input.
  \item Run Euclid on this reduced pair to get a short transformation matrix $M$.
  \item Apply $M$ to the \emph{full} inputs.
  \item If needed, do one exact Euclid correction step.
  \item Recurse on the now smaller pair.
\end{enumerate}
That is the entire idea: \emph{predict with high parts, apply exactly on full parts, recurse}.

\paragraph{Relation to the polynomial case.}
The same structure works for polynomials:
replace bit-length by degree, integer division by polynomial division,
and ``high bits'' by ``high-degree terms''.
This is why Half-GCD is a standard building block in fast polynomial
extended GCD, inversion modulo a polynomial, and related algorithms.

\begin{lstlisting}[language=Python,
  caption={Pedagogical Half-GCD (integer version, matrix-batched Euclid)}
]
def _mat_mul(A, B):
  return (
    (A[0][0] * B[0][0] + A[0][1] * B[1][0],
     A[0][0] * B[0][1] + A[0][1] * B[1][1]),
    (A[1][0] * B[0][0] + A[1][1] * B[1][0],
     A[1][0] * B[0][1] + A[1][1] * B[1][1]),
  )


def _apply_mat(M, a, b):
  c = M[0][0] * a + M[0][1] * b
  d = M[1][0] * a + M[1][1] * b
  return c, d


def _hgcd_step(a, b):
  """One half-gcd reduction step using top halves only."""
  n = max(a.bit_length(), b.bit_length())
  m = n // 2
  ah, bh = a >> m, b >> m

  M = ((1, 0), (0, 1))
  while bh and bh.bit_length() > m:
    q = ah // bh
    ah, bh = bh, ah - q * bh
    T = ((0, 1), (1, -q))
    M = _mat_mul(T, M)

  return _apply_mat(M, a, b)


def half_gcd(a, b):
  """Returns gcd(a,b). Correct, compact, and easy to read."""
  a, b = abs(a), abs(b)
  if b == 0:
    return a

  while b:
    if b.bit_length() < 64:
      while b:
        a, b = b, a % b
      return a

    c, d = _hgcd_step(a, b)

    if d == 0:
      return abs(c)

    # Safety fallback: if no progress, do one exact Euclid step.
    if abs(d) >= abs(b):
      a, b = b, a % b
    else:
      a, b = abs(c), abs(d)

  return a
\end{lstlisting}

\paragraph{Reading the code.}
The function \texttt{\_hgcd\_step} computes one ``batched Euclid'' move from high bits.
Then \texttt{half\_gcd} repeatedly applies such moves.
The fallback line keeps the implementation robust and easy to trust:
if a batched move does not shrink enough, do one ordinary Euclid step.
So this stays correct while still benefiting from larger jumps.

\section{Lagrange's Interpolation} \label{lagrange-interpolation}
Given a set of points $(x_0,y_0),\dots,(x_n,y_n)$, 
we can find a unique polynomial of degree-bound $n$ that passes through all of them.

The output of the algorithm is a polynomial in coefficient form.
This is a generalisation of inverse NTT for evaluation on 
arbitrary points.

Naively computing the Lagrange polynomial costs $O(n^3)$, but it can be done in $O(n^2)$ time by carefully reusing intermediate computations.

The key insight is that any polynomial of degree at most $n$ passing through $n+1$ points is unique. It therefore suffices to construct a polynomial of degree at most $n$ that passes through the given $n+1$ points.

\[\mathcal{L}(x)=\sum_{i=0}^{n}{L_i(x)\cdot y_i}\]

where \[L_i(x) =
\begin{cases}
  1 & \text{if } x=x_i \\
  0 & \text{otherwise}
\end{cases}\]

\paragraph{Constructing \(L_i(x)\)}

We need a polynomial of degree at most $n$ that equals $1$ at $x_i$ and $0$ at every other point $x_j$ ($j \neq i$).
The zero condition is satisfied by including a factor $(x - x_j)$ for each $j \neq i$, giving a product of $n$ terms.
Dividing by the constant $\prod_{j \neq i}(x_i - x_j)$ normalises the polynomial so that it evaluates to $1$ at $x_i$:

\[L_i(x) = \prod_{j=0,\, j \neq i}^{n} \frac{x - x_j}{x_i - x_j}.\]

\begin{notice}{Information}
  Lagrange's insight is that he did not need to know what the target polynomial looks like.
  He only needed to know that a polynomial of degree at most $n$ passing through $n+1$ points is unique
  (see~\ref{polynomials:point-val-representation}).
  It therefore suffices to construct \emph{any} polynomial of degree at most $n$ that passes through the same $n+1$ points.
  Uniqueness guarantees that this constructed polynomial must equal the target.
\end{notice}


\paragraph{Algorithm}

The naive approach computes each $L_i(x)$ independently, repeating work shared across all basis polynomials.
The efficient algorithm avoids this by factoring out the common nodal polynomial
\(m(x) = \prod_{i=0}^{n}(x - x_i)\),
so that \(L_i(x) = m(x)\,/\,[(x - x_i)\cdot c_i]\) where \(c_i = \prod_{j \neq i}(x_i - x_j)\).

\medskip
\noindent\textbf{Step 1: Build $m(x)$.}
Multiply in the linear factors one by one:
\[m^{(0)} = (x - x_0),\qquad m^{(k)} = m^{(k-1)}\cdot(x - x_k),\quad k=1,\dots,n.\]
Each multiplication extends a degree-$k$ polynomial by one factor in $O(k)$ coefficient operations, so the total cost is $O(n^2)$.

\medskip
\noindent\textbf{Step 2: Accumulate the interpolant.}
For each $i = 0, \dots, n$:
\begin{enumerate}
  \item Deflate: compute $m_i = m\,/\,(x - x_i)$ by synthetic division in $O(n)$.
  \item Normalise: evaluate $c_i = m_i(x_i) = \prod_{j \neq i}(x_i - x_j)$ in $O(n)$.
  \item Accumulate: $\mathcal{L} \mathrel{+}= (y_i / c_i)\cdot m_i$.
\end{enumerate}
Step 2 runs $n+1$ iterations, each costing $O(n)$, for a total of $O(n^2)$.

The output is $\mathcal{L}(x)$ in coefficient form, computed in $O(n^2)$ overall.

% \section{Reed-Solomon}
% It is fairly simple, but effective algorithm to encode words and be able to deal with erasurs.
% \subsection{Reed-Solomon for erasures}
% The Reed-Solomon algorithm is a systematic error-correcting code that can correct multiple erasures.
% It is based on polynomial arithmetic over a finite field, and it is used in many applications.

% The algorithm works as follows:
% encode the message in a polynomial of degree bound $k-1$,
% and then evaluate the polynomial at $n$ distinct points.

% To decode the message, we need to find the polynomial that interpolates the points we received.
% Given any $k$ points, once can find a unique polynomial of degree $k-1$ that passes through them
% using the Lagrange interpolation formula.


% \paragraph{Using NTT for Reed-Solomon} 
% The Reed-Solomon algorithm can be accelerated by employing the NTT algorithm (see \autoref{ntt}).
% Instead of utilizing any arbitrary $n$ distinct points, we can leverage the $n$-th 
% roots of unity (provided they exist within the finite field under consideration).
% Now we can encode messages faster.

% Decoding can still be done using the Lagrange interpolation formula.
% \todo{How to improve decoding speed using NTT!} 


% \subsection{Reed-Solomon for errors}
% \todo[inline]{This section is not complete}
% The Reed-Solomon algorithm can also correct data corruption.


% decoding can be done using the Berlekamp-Welch algorithm, which is a generalization of the Euclidean algorithm.
% It can be used to find the error locations and values, and then correct them.




% \subsection{Naive solution}
% assuming we decide we want our words to be of size $k=4$, and we add additional two redundancies $n=6$. 
% Thus to send it we'll do the following: given a word vector, we'll treat it as coefficients values of a 
% polynomial of degree $k-1$ (because we have k coefficients its a k-1 degree poly), then to encode it, we'll evaluate it on the following $x$ values: $(-1,0,1,2,\dots n)$.


% Now given the 6 values that we received over the network we can decode it as follows: 
% We go over every group of 4 points, and try to reconstruct the polynomial. 

% the polynomial we've reconstructed the most is the correct one.

% \subsection{Systematic Reed-Solomon}
% Assuming we want to be able to correct up to $s$ values, we'd need at least $2s$ redundant symbols.
% So if $k=4$, and $n=6$, we can only recover a single byte.

% It won't work for larger values, like $k=223$, $n=255$ which is a popular value for reed-solomon.
% (Meaning we add 32 more symbols to our message, so we can repair 16 symbols in total).

% The previous algorithm will work, but we'd need to send different values then 
% the original message. It would be best to send $Message||\text{addition code word}$.
% that way,we might be able to avoid reconstruction if we can find quickly that there was no error.

% this is called a Systematic code. 
