\chapter{Probability}

\section{Probability recap}
\begin{defn}
Probability space (merchav midgam)

a set $\Omega$ and a probability measure $p_{r}$. 
\end{defn}


\subsubsection*{Simplest case:}

Finite discrete probability space $\left|\Omega\right|<\infty$ individual
element $\omega\in\Omega$ we associate non negative number $p_{r}\left(\omega\right)\ge0$
s.t $\sum\limits _{\omega\in\Omega}p_{r}\left(\omega\right)=1$.
\begin{defn}
Events (meora)

We call $A\subseteq\Omega$ an event. The probability of an event
is: $p_{r}\left(A\right)=\sum\limits _{\omega\in A}p_{r}\left(\omega\right)$. 
\end{defn}


\subsubsection*{Obvious extension: }

$\left|\Omega\right|=\aleph_{0}$ for example $\Omega=\mathbb{N}$.
$\Omega=\left\{ x_{1},x_{2},\dots\right\} $ and $p_{1}.p_{2}\dots\ge0,\sum p_{i}=1$.

\subsubsection*{Last case:}

$\Omega=\left[0,1\right],0<\alpha<\beta<1,A=\left[\alpha,\beta\right]$.
$p_{r}\left(A\right)=\beta-\alpha$
\begin{defn}
Independent events (meoraot bilti tloolim)

Let $\left(\Omega,p_{r}\right)$ be a probability space\c{ }and let
$A,b\subseteq\Omega$ be two events. We say that events $A$ and $B$
are independent if $p_{r}\left(A\cap B\right)=p_{r}\left(A\right)\cdot p_{r}\left(B\right)$
\end{defn}

\begin{example}
intuition for random variables

a fair dice and a fair coin are being tossed. $\Omega=\left[6\right]\times\left\{ \mathbb{H},\mathbb{T}\right\} $,
$\forall\omega\in\Omega,p_{r}\left(\omega\right)=\frac{1}{12}$ (uniform
distribution). We treat the dice and the coin as independent. Thus
$p_{r}\left(\text{dice is even}\right)=p_{r}\left(A\right)=\frac{1}{2}$,
$p_{r}\left(\text{coin is on }H\right)=p_{r}\left(B\right)=\frac{1}{2}$.
$p_{r}\left(\text{dice is even}\cap\text{coin is on }H\right)=p_{r}\left(A\cap B\right)=p_{r}\left(A\right)\cdot p_{r}\left(B\right)=\frac{1}{4}$.
Now we can inspect all possible cases of $\left[6\right]\times\left\{ \mathbb{H},\mathbb{T}\right\} $
and find the event. $A\cap B=\left\{ \left(2,H\right),\left(4,H\right),\left(6,H\right)\right\} $
thus $p_{r}\left(A\cap B\right)=\frac{3}{\left|\Omega\right|}=\frac{3}{12}=\frac{1}{4}$
\end{example}

\begin{defn}
Conditioning (hatnaya)
\begin{defn}
conditioning on the event $A$ is defined as follows: $p_{r}\left(B|A\right)=\frac{p_{r}\left(A\cap B\right)}{p_{r}\left(A\right)}$.
In case of independence between the events it is simple to deduce
that $p_{r}\left(B|A\right)=p_{r}\left(B\right)$: 
\[
p_{r}\left(B|A\right)=\frac{p_{r}\left(A\cap B\right)}{p_{r}\left(A\right)}=\frac{p_{r}\left(A\right)\cdot p_{r}\left(B\right)}{p_{r}\left(A\right)}=p_{r}\left(B\right)
\]
.
\end{defn}

\end{defn}

\begin{rem}
We can think of conditioning as looking at event $A$ as our probability
space thus we normalize according to it. $A$ happened what is the
probability that $B$ will happen too? so we do a conjunction between
the two and normalize by A to get the new probability.
\end{rem}

\begin{defn}
Random Variable: Let $\left(\Omega,p_{r}\right)$ be a probability
space and $X:\Omega\to\mathbb{R}$ is called a real random variable.
In general it is very typical to not care about where $X$ maps $\omega\in\Omega$.
In most cases we care about the event that is tied to the random variable. 
\end{defn}

\begin{example}
Random Variable: 

$X:\Omega\to\mathbb{R}$, $\left[\alpha,\beta\right]\subseteq\mathbb{R}$,
let $A\subseteq\Omega$ be the event $\left\{ \omega\in\Omega|X\left(\omega\right)\in\left[\alpha,\beta\right]\right\} $.
Using random variable we this way allows us to inquire important information
about real life events. Each $\omega\in\Omega$ in this example can
be a person and $X$ can map from a person to it's salary. 
\end{example}

\begin{defn}
Random variable independency 

Let $X,Y:\Omega\to S$ be Random Variables (R.V) we say that $X,Y$
are independent if $\forall r,\sigma\in S$ the events $\left\{ \omega|X\left(\omega\right)=r\right\} ,\left\{ \omega|Y\left(\omega\right)=\sigma\right\} $
are independent events.
\end{defn}

\begin{defn}
Indicator 

$X:\Omega\to\left\{ 0,1\right\} $ is called an ``indicator random
variable''.
\end{defn}

When we are working with random variables we might face questions
like the following: What is the average grade of $X$
\begin{defn}
Expectation (Tochelet)

Denote $\mathbb{E}\left[x\right]=\sum\limits _{\omega\in\Omega}p_{r}\left(\omega\right)\cdot X\left(\omega\right)$
the expectation of $X$.
\end{defn}

\begin{claim}
Linearity of expectation (with no proof)

$\forall X,Y:\Omega\to\mathbb{R},\forall a,b\in\mathbb{R},\mathbb{E}\left[aX+bY\right]=a\mathbb{E}\left[X\right]+b\mathbb{E}\left[Y\right]$
\end{claim}

Aside from the average (expectation) of a R.V we might also wish to
know how \uline{concentrated} it is. how can we find how far are
all the values $p_{1}a_{1}+\dots+p_{n}a_{n}=\mathbb{E}\left[X\right]$
from the average of $X$?

\paragraph*{First try}

$\mu:=\mathbb{E}\left[X\right]=\sum p_{i}a_{i}$ , now to calculate
how far each value (with it's appropriate size factor):$\sum p_{i}\left(a_{i}-\mu\right)=\sum\left(p_{i}a_{i}\right)-1\cdot\sum p_{i}a_{i}=0$
\\
Didn'\'{t} go as wanted.

A working definition would be $Var\left[X\right]=\sum p_{i}\left(a_{i}-\mu\right)^{2}$
\begin{defn}
Variance of a random variable

$Var\left[X\right]=\sum p_{i}\left(a_{i}-\mu\right)^{2}=\mathbb{E}\left[\left(X-\mu\right)^{2}\right]=\mathbb{E}\left[\left(X-\mathbb{E}\left[X\right]\right)^{2}\right]$
Let as simplify this equation:
\begin{align*}
\mathbb{E}\left[\left(X-\mathbb{E}\left[X\right]\right)^{2}\right] & =\mathbb{E}\left[X^{2}-\underbrace{2\mathbb{E}\left[X\right]}_{scalar}\cdot X+\mathbb{E}\left[X\right]^{2}\right]=\\
\text{Linearity} & =\mathbb{E}\left[X^{2}\right]-2\cdot\mathbb{E}\left[X\right]\cdot\mathbb{E}\left[X\right]+\mathbb{E}\left[X\right]^{2}=\\
 & =\mathbb{E}\left[X^{2}\right]-2\mathbb{E}\left[X\right]^{2}+\mathbb{E}\left[X\right]^{2}=\\
 & =\mathbb{E}\left[X^{2}\right]-\mathbb{E}\left[X\right]^{2}
\end{align*}
\end{defn}

\begin{thm}[Markov inequality]
 If $X$ is a non-negative R.V and $c>1$ then $p_{r}\left(x\ge c\cdot\mathbb{E}\left[X\right]\right)\le\frac{1}{c}$
\end{thm}

\begin{theorem}[Chebyshev inequality]
    \label{probabilty:chebyshev}
 If $X$ is a real R.V then $p_{r}\left(\left|X-\mathbb{E}\left[X\right]\right|\ge c\cdot\sigma\right)\le\frac{1}{c^{2}}$
where $var\left[X\right]=\sigma^{2}$ .

In general we can view Chebyshev's inequality as a way to see how
far $X$ is from It's expectation by units of it's standard deviation
($\sigma=\sqrt{Var\left[X\right]}$).

The intuition of proving that claim is to apply Markov's inequality
over $Y=\left(X-\mathbb{E}\left[X\right]\right)^{2}$ and understanding
that $p_{r}\left(\left(x-\mathbb{E}\left[X\right]\right)^{2}\ge c^{2}\sigma^{2}\right)=p_{r}\left(\left|X-\mathbb{E}\left[X\right]\right|\ge c\cdot\sigma\right)$.

\end{theorem}

\begin{xca}
If you pick at random a permutation $\pi:\left[n\right]\to\left[n\right]\in S_{n}$
How many fixed points does it have?

In order to answer this question we will create a random variable
that counts the number of fixed points in a given permutation. afterward
we will search it's expectation. Let $X\left(\pi\right)=\sum\limits _{i=1}^{n}X_{i}\left(\pi\right)$
where $X_{i}=\begin{cases}
1 & \pi\text{ has a fixed point at }i\\
0 & otherwise
\end{cases}$ . Thus $X$ is a counter for fixed points in $\pi$. Let calculate
the expectation of $X$: $\mathbb{E}\left[X\right]=\mathbb{E}\left[\sum_{i}X_{i}\right]$
and from linearity of expectation $\mathbb{E}\left[X\right]=\sum\limits _{i=1}^{n}\mathbb{E}\left[X_{i}\right]$.
Now we need to find the expectation of the indicator. $\mathbb{E}\left[X_{i}\right]=1\cdot p_{r}\left(\text{there is a fixed point at }i\right)+0\cdot p_{r}\left(otherwise\right)$.
$p_{r}\left(\text{there is a fixed point at }i\right)=\frac{\left|\left\{ \pi|\pi\text{ has a fixed point at }i\right\} \right|}{n!}.$
The value of $\left|\left\{ \pi|\pi\text{ has a fixed point at }i\right\} \right|$
is $\left(n-1\right)!$ because we fixed one position and then we
still have $n-1$ elements to choose for them a position, therefore
$\left(n-1\right)!$ possible permutations. As a result $p_{r}\left(\text{there is a fixed point at }i\right)=\frac{\left(n-1\right)!}{n!}=\frac{1}{n}$.
Which in turn means that $\mathbb{E}\left[X\right]=\sum\limits _{i=1}^{n}\frac{1}{n}=1$.
The answer to our question is 1.
\end{xca}

\begin{rem}
The above exercise is known as the disoriented secretary.
\end{rem}


\section{Ramsey's Theorem and the probabilistic method}

First lets define $R\left(k,l\right)$: the minimum number of vertices
in a graph that is needed in order to make either $k$ size clique
that is coloured blue on it's edges or $l$ red clique. (clique and
anti-clique)
\begin{example}
What is the number of people needed to be invited to a party in order
to be sure that either 3 know each other or 3 do not know each other?
$R\left(3,3\right)=6$.
\end{example}

\begin{thm}[Erd�s-Szekeres 35]
 $R\left(k,l\right)\le{k+l-2 \choose k-1}$. The theorem is without
proof.

For $R\left(k,k\right)\le{k+k-2 \choose k-1}=\binom{2k-2}{k-1}\le2^{2k}=n$,
it implies that in every graph with $n$ vertices there exists a clique
or an anti-clique of size $\ge\frac{1}{2}\log_{2}n$ (because $\frac{1}{2}\log_{2}n=\frac{1}{2}\log_{2}\left(2^{2k}\right)=k\log_{2}2=k$). 

\end{thm}

\begin{rem}
\label{probRVreualsZero}if $X$ is a R.V that is non negative then
$p\left[X\ge c\cdot\mathbb{E}\left[X\right]\right]\le\frac{1}{c}$.
If X takes values in $\N\cup\left\{ 0\right\} $ and $\mathbb{E}\left[X\right]=\epsilon$
then $P\left[X=0\right]\ge1-\epsilon$. (from Markov) 
\end{rem}

\begin{rem}
Chebyshev inequality: $p\left[X=0\right]\le\frac{Var\left[X\right]}{\mathbb{E}\left[X\right]^{2}}$
(proved in problem set 1).
\end{rem}

\begin{corollary}
$\forall N$ every $N$-vertex graph must have a clique or an anti-clique
on $\ge\frac{1}{2}\log_{2}N$ vertices.
\end{corollary}

\begin{thm}[Erdos '47]
 $\exists N$-vertex graph with no clique nor anti-clique $>2\log_{2}N$.
In fact this is true for almost all graphs.
\end{thm}

\begin{proof}
$\Omega_{n}=\left\{ \text{all graphs on }V=\left\{ 1,2,\dots N\right\} \right\} $,
there is a uniform distribution; $\forall G\in\Omega_{N},p_{r}\left[G\right]=\frac{1}{2^{\binom{n}{2}}}$.
\begin{rem}
the number of all possible graphs is $2^{\binom{n}{2}}$ because for
each edge we have two choices and we have $\binom{n}{2}$ edges.
\end{rem}

This probability space is named $G\left(n,\frac{1}{2}\right)$. $k$
is a parameter, $X,Y:\Omega_{n}\to\N\cup\left\{ 0\right\} $. $X\left(G\right)=\text{Number of k-cliques in }G$.
Y is the same but for anti-cliques.
\begin{proposition}
Suppose that $\mathbb{E}\left[X+Y\right]=o\left(1\right)$, where
$o\left(1\right)$ is something that goes to zero when $N\to\infty$.
If this happens then we can conclude that $p_{r}\left[X+Y=0\right]=1-o\left(1\right)$
(remember remark \ref{probRVreualsZero}). Well that means that $X=Y=0$
and that most $N$ - vertex graphs have no k-clique and no k-anti-clique.
So we want to find that $\mathbb{E}\left[X+Y\right]=o\left(1\right)$.
\begin{rem}
Please note that $\mathbb{E}\left[X+Y\right]=2\cdot\mathbb{E}\left[X\right]$
because they are the same R.V but for different colour.
\end{rem}

\end{proposition}

Lets start: We define $X=\sum\limits _{\begin{array}{c}
S\subseteq V\\
\left|S\right|=k
\end{array}}X_{S}$ where $S$ are groups of $k$ vertices from $V$, where
\[
X_{s}=\begin{cases}
1 & \text{S is a clique in G}\\
0 & \text{otherwise}
\end{cases}
\]
 now lets find the expectation:
\begin{align*}
\mathbb{E}\left[X\right] & =\sum\limits _{S}\mathbb{E}\left[X_{S}\right]=\\
 & =\binom{n}{k}p_{r}\left[\underbrace{\left\{ 1,\dots k\right\} }_{vertices}\text{ is a clique}\right]=\\
 & =\binom{n}{k}\cdot\left(\frac{1}{2}\right)^{\binom{k}{2}}
\end{align*}
notice that $\left(\frac{1}{2}\right)^{\binom{k}{2}}$ is the probability
to form an edge $\binom{k}{2}$ times (between all vertices). Lets
inspect the expression above again:
\[
=\frac{\binom{n}{k}}{\left(2\right)^{\binom{k}{2}}}
\]
We need to show that $\binom{n}{k}\cdot\left(\frac{1}{2}\right)^{\bin k2}=o\left(1\right)$.
Notice that $\binom{n}{k}<n^{k}<2^{\frac{k\left(k-1\right)}{2}}\Rightarrow\binom{N}{k}\ll2^{\binom{k}{2}}$.
Consequently the expectation strives to zero as $k$ tends to infinity.
Lets choose $k$ in respect to $N$: 
\begin{align*}
N^{k} & <2^{\frac{k\left(k-1\right)}{2}}\\
N & <2^{\frac{k-1}{2}}\\
\log_{2}N & <\frac{k-1}{2}\\
1+2\cdot\log_{2}N & <k
\end{align*}
 for $k>1+2\log_{n}N$ it holds that $\mathbb{E}\left[X+Y\right]=o\left(1\right)$
and then $p_{r}\left[X+Y=0\right]=1-o\left(1\right)$ as wanted.
\end{proof}
\begin{example}
We will inspect another view of this method: What is the probability
that $G\sim G\left(n,p\right)$ contains a $K_{4}$? ($k_{4}$is a
clique of 4).
\begin{rem}
$G\left(n,p\right);p_{r}\left[G\right]=p^{\left|E\left(G\right)\right|}\cdot\left(1-p\right)^{\binom{n}{2}-\left|E\left(G\right)\right|}$.
Explanation: $p^{\left|E\left(G\right)\right|}$ means the probability
to get the edges. $\left(1-p\right)^{\binom{n}{2}-\left|E\left(G\right)\right|}$
means to not get an edge number of possible edges in the graph minus
the number of edges we already have.
\end{rem}

\begin{thm}
$p=n^{-\frac{2}{3}}$is the threshold probability for the occurrences
of $k_{4}$ in a $G\left(n,p\right)$ graph.
\begin{enumerate}
\item for $p=o\left(n^{-\frac{2}{3}}\right)$ i.e $p\ll n^{-\frac{2}{3}}$
``almost surely'' G contains no $k_{4}$.
\item for $p=\omega\left(n^{-\frac{2}{3}}\right)$ i.e $p\gg n^{-\frac{2}{3}}$
``almost surely'' G contains $k_{4}$.
\end{enumerate}
\begin{rem}
``almost surely'' means: when we sample with p from the graph ,
and the probability strives to zero as $n\to\infty$.
\end{rem}

\end{thm}

$X:G\left(n,p\right)\to\N\cup\left\{ 0\right\} $, $X\left(G\right)=\text{num }k_{4}\text{s in }G$.
$X=\sum\limits _{\left|S\right|=4}X_{s}$ where $X_{s}=\begin{cases}
1 & \text{S is a 4-clique in G}\\
0 & o/w
\end{cases}$. 
\end{example}

\begin{casenv}
\item $p=o\left(n^{-\frac{2}{3}}\right)$ and as a result 
\begin{align*}
\mathbb{E}\left[X\right] & =\sum\mathbb{E}\left[X_{s}\right]=\binom{n}{4}\cdot p_{r}\left[\left\{ 1,2,3,4\right\} \text{ is a clique}\right]=\\
 & =\binom{n}{4}p^{\binom{4}{2}}=\binom{n}{4}p^{6}\le n^{4}p^{6}\\
 & =o\left(1\right)
\end{align*}
As a result we recall $p_{r}\left[X=0\right]=1-\E X=1-o\left(1\right)$
as wanted.
\item $p=\omega\left(n^{-\frac{2}{3}}\right)$: recall that if $X:\Omega\to\N\cup\left\{ 0,1,2,\dots\right\} $
then $p_{r}\left[X=0\right]\le\frac{var\left[X\right]}{\mathbb{E}\left[X\right]^{2}}$.
We want to show that $var\left[X\right]\ll\mathbb{E}\left[X\right]^{2}$.
$X$ is the same, $\mathbb{E}\left[X\right]$ is the same. Lets calculate
the variance: 
\begin{align*}
var\left[X\right] & =\mathbb{E}\left[X^{2}\right]-\mathbb{E}\left[X\right]^{2}=\\
 & =\sum_{\left|S\right|=4}\left(\mathbb{E}\left[X_{s}^{2}\right]-\mathbb{E}\left[X_{s}\right]^{2}\right)+2\cdot\sum_{\begin{array}{c}
S\ne T\\
\left|S\right|=\left|T\right|=4
\end{array}}\left(\underbrace{\mathbb{E}\left[X_{s}\cdot X_{T}\right]-\mathbb{E}\left[X_{s}\right]\cdot\mathbb{E}\left[X_{T}\right]}_{cov\left(X_{s},X_{t}\right)}\right)=\\
 & =\sum_{\left|S\right|=4}\left(\mathbb{E}\left[X_{s}\right]-\mathbb{E}\left[X_{s}\right]^{2}\right)+2\cdot\sum_{\begin{array}{c}
S\ne T\\
\left|S\right|=\left|T\right|=4\\
\left|S\cap T\right|\in\left\{ 2,3\right\} 
\end{array}}\left(\underbrace{\mathbb{E}\left[X_{s}\cdot X_{T}\right]-\mathbb{E}\left[X_{s}\right]\cdot\mathbb{E}\left[X_{T}\right]}_{cov\left(X_{s},X_{t}\right)}\right)
\end{align*}
First let us handle the Cov summation: Notice that there are 2 cases
to handle $\left|S\cap T\right|=2$ and $\left|S\cap T\right|=3$,
that is because the R.V are either independent when $\left|S\cap T\right|=1$
or they are disjointed. when $\left|S\cap T\right|=4$ they are the
same variable hence not in the summation.
\begin{rem}
Note that we indeed compute only where the covariance is no\'{t} zero,
but we can enlarge the value of $Var\left[X\right]$ and make our
life easier by simply computing 
\begin{align*}
\sum\limits _{\left|S\cap T\right|\in\left\{ 2,3\right\} }\mathbb{E}\left[X_{s}\cdot X_{T}\right]-\mathbb{E}\left[X_{s}\right]\cdot\mathbb{E}\left[X_{T}\right]\\
<\sum\limits _{\left|S\cap T\right|\in\left\{ 2,3\right\} }\mathbb{E}\left[X_{s}\cdot X_{t}\right]
\end{align*}
 so we will address only the values in $\sum\limits _{\left|S\cap T\right|\in\left\{ 2,3\right\} }\mathbb{E}\left[X_{s}\cdot X_{t}\right]$.
\end{rem}

\begin{casenv}
\item 
\begin{align*}
\sum\limits _{\left|S\cap T\right|=2}\mathbb{E}\left[X_{s}\cdot X_{t}\right] & =\\
\underbrace{\binom{n}{4}}_{\left(1\right)}\cdot\underbrace{\binom{4}{2}}_{\left(2\right)}\cdot\underbrace{\binom{n-4}{2}}_{\left(3\right)}\cdot\underbrace{p^{11}}_{\left(4\right)}
\end{align*}
 explanation:
\begin{enumerate}
\item number of possibilities to choose the set $S$ (choose 4 vertices
out of the n possible vertices).
\item the possibilities to choose two vertices off of $S$ to be in the
intersection $T\cap S$.
\item all other possibilities for the 2 other vertices of $T$.
\item 11 is the number of edges we needed to to be connected. 
\end{enumerate}
\item 
\[
\sum\limits _{\left|S\cap T\right|=3}\mathbb{E}\left[X_{s}\cdot X_{t}\right]=\binom{n}{4}\cdot\binom{4}{3}\cdot\binom{n-4}{1}\cdot p^{9}
\]
 The explanation is similar to case $i$.
\end{casenv}
\begin{align*}
var\left[X\right] & \le\\
 & \mathbb{E}\left[X\right]+\underbrace{\Theta\left(n^{6}p^{11}\right)+\Theta\left(n^{5}p^{9}\right)}_{\begin{array}{c}
\text{we removed the array because }\\
\text{it is encapsulated in the Thetas}
\end{array}}\\
 & <\mathbb{E}\left[X\right]^{2}=\left(n^{4}p^{6}\right)^{2}=n^{8}p^{12}
\end{align*}
 $var\left[X\right]\ll\mathbb{E}\left[X\right]^{2}$ which means that
$p_{r}\left[X=0\right]\le\frac{var\left[X\right]}{\mathbb{E}\left[X\right]^{2}}\to0$
We got what we wanted: $X\ge1$ almost surely.\\

\end{casenv}
%

\section{Balls and bins}
\begin{fact}
$1+x\le e^{x}$ 
\end{fact}

\begin{fact}
for a small $x$: $1+x\ge e^{x}-O\left(x^{2}\right)$ 
\end{fact}


\subsection{Birthday paradox}

When we throw multiple balls into multiple bins ($n$ bins) we expect
to see the first collision (a second ball into an occupied bin) in
our $\sim\sqrt{n}$ throw. $p_{r}\left[\text{no collision}\right]=1\cdot\left(1-\frac{1}{n}\right)\cdots\left(1-\frac{k-1}{n}\right)$
that is the first ball doesn't have a collision with probability 1\c{ }the
second doesn't have a collision with probability $\left(1-\frac{1}{n}\right)$
because to hit the occupied bin is $\frac{1}{n}$ and that is its
complimentary event, and so on. 
\begin{align*}
p_{r}\left[\text{no collision}\right] & =1\cdot\left(1-\frac{1}{n}\right)\cdots\left(1-\frac{k-1}{n}\right)=\\
\left(\text{fact 32.}\right) & \le1\cdot e^{-\frac{1}{n}}\cdots e^{-\frac{k-1}{n}}=\\
 & =e^{-\left(\frac{1}{n}\sum_{i=1}^{k-1}i\right)}=e^{-\frac{1}{n}\left(\frac{k\left(k-1\right)}{2}\right)}=\\
 & =e^{-\frac{k\left(k-1\right)}{2n}}
\end{align*}
 We can see from the above expression that if $k=o\left(\sqrt{n}\right)$
then $e^{-\frac{k\left(k-1\right)}{2n}}$ the above expression as
n tends to infinity get closer to 1 because $k$ is a lot smaller
than $\sqrt{n}$ and as a result $\limn{e^{\frac{-k^{2}-k}{2n}}}=e^{0}$.
On the other hand $k=\omega\left(\sqrt{n}\right)$ $e^{-\frac{k\left(k-1\right)}{2n}}=o\left(1\right)$
. So we conclude that after $\sqrt{n}$ throws it is quite certain
that there will be collisions.
\begin{example}
Suppose we pick two random subsets $A,B$ of $\left\{ 1,\dots,n\right\} $
of cardinals (size) $k,l$. How likely is it that the two sets are
disjoint? 

Without the loss of generality (WLOG): $A=\left\{ 1\dots k\right\} $
and B is chosen uniformly at random. $p_{r}\left[A\cap B=\emptyset\right]=\frac{\binom{n-k}{l}}{\binom{n}{l}}$
. ($\binom{n-k}{l}$ is because we want to chose $l$ elements that
do not collide with elements from $A$, $\binom{n}{l}$ is all possible
sets of size $l$ - our probability space). 
\begin{align*}
\p{A\cap B=\emptyset} & =\frac{n-k}{n}\cdot\frac{n-k-1}{n-1}\cdot\dots\cdot\frac{n-k-l+1}{n-l+1}=\\
 & =\left(1-\frac{k}{n}\right)\cdot\left(1-\frac{k}{n-1}\right)\cdots\left(1-\frac{k}{n-l+1}\right)\cong\\
 & \cong e^{-\left(\frac{k}{n}+\frac{k}{n-1}+\dots+\frac{k}{n-l+1}\right)}\cong e^{-\frac{kl}{n}}
\end{align*}
\begin{lyxgreyedout}
\begin{align*}
p_{r}\left[A\cap B=\emptyset\right] & =\frac{\binom{n-k}{l}}{\binom{n}{l}}=\frac{\frac{\left(n-k\right)!}{l!\left(n-k-l\right)!}}{\frac{n!}{l!\left(n-l\right)!}}=\\
 & =\frac{\frac{\left(n-k\right)\left(n-k-1\right)\dots\left(n-k-l\right)\left(n-k-l-1\right)\dots}{l!\left(n-k-l\right)\left(n-k-l-1\right)\dots}}{\frac{n\left(n-1\right)\dots\left(n-l\right)\left(n-l-1\right)\dots}{l!\left(n-l\right)\left(n-l-1\right)\dots}}=\\
 & =\frac{\frac{\left(n-k\right)\left(n-k-1\right)\dots\left(n-k-l+1\right)}{l!}}{\frac{n\left(n-1\right)\dots\left(n-l\right)\left(n-l+1\right)}{l!}}=\\
 & =\frac{\left(n-k\right)\left(n-k-1\right)\dots\left(n-k-l+1\right)}{n\left(n-1\right)\dots\left(n-l\right)\left(n-l+1\right)}=\\
 & =\frac{n-k}{n}\cdot\frac{n-k-1}{n-1}\cdots\frac{n-k-l+1}{n-l+1}=\\
 & =\left(1-\frac{k}{n}\right)\left(1-\frac{k}{n-1}\right)\cdots\left(1-\frac{k}{n-l+1}\right)=\\
 & \backsimeq e^{-\left(\frac{k}{n}+\frac{k}{n-1}+\dots+\frac{k}{n-l+1}\right)}
\end{align*}
 please notice that $e^{-\frac{k}{n-1}}\backsimeq e^{-\frac{k}{n}}$
hence 
\[
\backsimeq e^{-\left(\underbrace{\frac{k}{n}+\frac{k}{n}+\dots+\frac{k}{n}}_{l\text{ times}}\right)}=e^{-\frac{kl}{n}}
\]
\end{lyxgreyedout}
{} Now, if $kl=\omega\left(n\right)$ then asymptotic almost surely
$\left(a.a.s\right)$ the two sets intersect because $p_{r}\left[A\cap B=\emptyset\right]=e^{-\frac{kl}{n}}=o\left(1\right)$.
On the other hand if $kl=o\left(n\right)$ then $p_{r}\left[A\cap B=\emptyset\right]=e^{-\frac{kl}{n}}=1-o\left(1\right)$.
\end{example}


\subsection{The coupons collector}
\begin{claim}
It takes roughly $n\log n$ throws of $m$ balls into $n$ bins in
order to make sure that we have inserted at least one ball into every
bin. 
\end{claim}

\begin{proof}
Let $X$ be the R.V that counts the number of empty bins. $X=\sum\limits _{i=1}^{n}X_{i}$
where $X_{i}$ is an indicator R.V for the event ``the $i$-th bin
is empty''. The random variable $X_{i}$ are identically distributed
and their expectation $\E{X_{i}}=\Pr\left[X_{i}=1\right]=\left(1-\frac{1}{n}\right)^{m}$
which means that after $m$ throws a specific bin is still empty.
By linearity of expectation\"{ }$\E X=\sum\limits _{i=1}^{n}\E{X_{i}}=n\cdot\left(1-\frac{1}{n}\right)^{m}\backsimeq n\cdot e^{\frac{-m}{n}}$. 
\begin{casenv}
\item if $m=\omega\left(n\log n\right)$ then $n\cdot e^{\frac{-m}{n}}=o\left(1\right)$
\item if $m=o\left(n\log n\right)$ then $n\cdot e^{\frac{-m}{n}}=\omega\left(1\right)$ 
\end{casenv}
\end{proof}
%
\begin{proof}
Another proof: define $T_{i}$ as the time between having seen $\left(i-1\right)$
bins to the first time bin $i-th$ is visited. this means that $T=\sum T_{i}$.
$\E T=\sum_{i=1}^{n}\E{T_{i}}$. 
\begin{rem}
$T_{i}$ is a geometrically distributed random variable $T_{i}\sim Geo\left(p\right)$.
$\E{T_{i}}=\frac{1}{p}$.
\end{rem}

thus $T_{i}\sim Geo\left(\frac{n-i+1}{n}\right)$, the probability
is defined as follows $n-i+1$ bins are currently empty, so we need
any one of them over the probability space which is $n$. Now we can
compute $\E T=\sum_{i=1}^{n}\E{T_{i}}=\sum_{i=1}^{n}\frac{n}{n-i+1}=n\sum_{i=1}^{n}\frac{1}{n-i+1}$
notice that $\sum_{i=1}^{n}\frac{1}{n-i+1}=\left(\frac{1}{n}+\frac{1}{n-1}+\dots\frac{1}{1}\right)=\sum\limits _{i=1}^{n}\frac{1}{i}$
and thus $\E T=n\sum\limits _{i=1}^{n}\frac{1}{i}=n\left(\log_{e}n+O\left(1\right)\right)$. 
\end{proof}
We would like to say that after exceeding $n\log n+O\left(n\right)$
throws it's almost certain that there is no empty bin. We cannot use
Markov's inequality because it isn't tight enough: 
\[
\Pr\left[T\ge10\cdot\E T\right]\le\frac{1}{10}\iff\Pr\left[T\ge10\cdot n\log n\right]\le\frac{1}{10}
\]
We can use Chebyshev's inequality
\[
\Pr\left[\left|T-\E T\right|\text{\ensuremath{\ge c\cdot\var T}}\right]\le\frac{1}{c^{2}}
\]
Notice that $\var T=\var{\sum T_{i}}$ and that the $T_{i}$ R.Vs
are independent %
\begin{lyxgreyedout}
Notice that each $T_{i}$ is independent because they count number
of tries given a static event. So we check for each$T_{i}$ how many
tries should we do when $i$ slots are already filled.%
\end{lyxgreyedout}
. 
\begin{rem}
A geometric distributed R.V $X$ with parameter $p$ has $\var X=\frac{1-p}{p^{2}}$.
\end{rem}

Therefore 
\[
\var T=\var{\sum T_{i}}=\sum\var{T_{i}}=\sum_{i=1}^{n}\left(\frac{n}{n-i+1}\right)^{2}=n^{2}\sum_{j=1}\frac{1}{j^{2}}\le\frac{\pi^{2}}{6}
\]
 using Chebyshev's inequality: 
\[
\Pr\left[T>n\ln n+cn\right]\le\frac{10}{c^{2}}
\]

\begin{lyxgreyedout}
Not sure whys that chebyshev%
\end{lyxgreyedout}
{} If we choose $c=\sqrt{\log n}$ we get that the probability tends
to 0.

\section{Random walks}

A drunkard takes a walk, the drunkard can either take a step to the
right or to the left. There are multiple questions that might interest
us, such as: 
\begin{enumerate}
\item How far typically will the drunkard reach? $\pm O\left(\sqrt{n}\right)$. 
\item How often will the drunkard return to the origin of the random walk?
$\Theta\left(\sqrt{n}\right)$.
\item We can ask the same questions above higher dimensions. 
\end{enumerate}
We can model the problem as having $N$ RVs $X_{1},\dots,X_{n}$ which
are i.i.d (independent identically distributed). $\p{X_{i}=1}=\p{X_{i}=-1}=\frac{1}{2}$,
$X:=\sum\limits _{i=1}^{N}X_{i}$ ($X$ represents our random walk
after $N$ steps). 
\begin{proposition}
$\p{X\ge a}\le e^{-\frac{a^{2}}{2N}}$ We will use this and prove
it afterward. 
\begin{example}
$\p{X>5\sqrt{N}}\le e^{-\frac{25N}{2N}}=e^{\frac{-25}{2}}$, using
that fact we get that $X$ is quite close to $\sqrt{N}$. In fact
the probability of $X$ being bigger than $\sqrt{N}$ or smaller than
$-\sqrt{N}$ tends to zero.
\end{example}

\end{proposition}

\begin{proof}
$\forall t>0$, $\p{X\ge a}=\p{e^{tX}\ge e^{ta}}$ (the events $e^{tX}\ge e^{ta}$
is identical to $X\ge a$). 
\begin{align*}
\p{\underbrace{e^{tX}}_{R.V}\ge e^{ta}} & \overset{\left(Markov\right)}{\le}\frac{1}{\left(\frac{e^{ta}}{\E{e^{tX}}}\right)}=e^{-ta}\E{e^{tX}}=\\
 & =e^{-ta}\E{e^{t\sum_{i=1}^{N}X_{i}}}=e^{-ta}\E{\prod_{i=1}^{N}e^{tX_{i}}}=\\
\left(independence\right) & =e^{-ta}\prod_{i=1}^{N}\E{e^{tX_{i}}}
\end{align*}
 We know that $X_{i}\left(\omega\right)\in\left\{ -1,1\right\} $,
as a result $\E{e^{tX_{i}}}=\frac{1}{2}\cdot\left(e^{t}+e^{-t}\right)$.
collecting everything up to now: $\p{X\ge a}\le e^{-ta}\left[\frac{e^{t}+e^{-t}}{2}\right]^{N}$.
\begin{claim}
$\forall t\ge0$, $\frac{e^{t}+e^{-t}}{2}\le e^{\frac{t^{2}}{2}}$.
\begin{proof}
$e^{u}=\sum\limits _{j=0}^{\infty}\frac{u^{j}}{j!}\Rightarrow$ 
\begin{align*}
e^{t} & =1+\frac{t}{1!}+\frac{t^{2}}{2!}+\dots\\
e^{-t} & =1+\frac{(-t)}{1!}+\frac{(-t)^{2}}{2!}\dots
\end{align*}
 Notice that the even powers are with the same sign, and the odd powers
cancel each other. we get that
\[
\frac{e^{t}+e^{-t}}{2}=\sum\limits _{j=0}^{\infty}\frac{t^{2j}}{\left(2j\right)!}\le\sum\limits _{j=0}^{\infty}\frac{\left(t^{2}\right)^{j}}{2^{j}\cdot j!}=\sum\limits _{j=0}^{\infty}\frac{\left(\frac{t^{2}}{2}\right)^{j}}{j!}=e^{\frac{t^{2}}{2}}
\]
 
\end{proof}
now lets continue 
\[
\p{X\ge a}\le e^{-ta}\cdot\left[\frac{e^{t}+e^{-t}}{2}\right]^{N}\le e^{-ta}\cdot e^{\frac{t^{2}N}{2}}=e^{\frac{t^{2}N}{2}-ta}
\]
 choose $t=\frac{a}{N}$ and we get 
\[
\p{X\ge a}\le e^{\frac{t^{2}N}{2}-ta}=e^{\frac{\left(\frac{a}{N}\right)^{2}N}{2}-\frac{a}{N}a}=e^{-\frac{a^{2}}{2N}}
\]

as wanted.
\end{claim}

\end{proof}
%
\begin{fact}
$\sum_{i=1}^{N}\binom{N}{i}=2^{n}$
\end{fact}

\begin{fact}
Stirling Formula: $k!\approx\left(\frac{k}{e}\right)^{k}\cdot\sqrt{2\pi k}$ 
\end{fact}

Assuming that $N$ is even $X=\sum\limits _{i=1}^{N}X_{i}$ then the
probability to go back to origin is $\p{X=0}=\frac{\binom{N}{N/2}}{2^{N}}$,
where $\binom{N}{N/2}$ is all possible choosing for $N/2$ steps
right and the the other $N/2$ steps are to the left, and $2^{N}$
are all possible sequence of steps over $N$ walks. The probability
after $N$ steps to be two steps to the right from the origin namely
$\p{X=2}$ is $\frac{\binom{N}{1+N/2}}{2^{N}}$, where $1+N/2$ because
we do $1+N/2$ steps to the right and $\left(N/2\right)-1$ steps
to the left.

A useful lemma to express with better ease expressions like above
$\left(\p{X=0}=\frac{\binom{N}{N/2}}{2^{N}}\right)$:
\begin{lemma}
$\forall0<\alpha<1$, $\bin N{\alpha N}=2^{1+o\left(1\right)}\cdot2^{NH\left(\alpha\right)}$
where $H\left(\alpha\right)=-\alpha\log_{2}\alpha-\left(1-\alpha\right)\log_{2}\left(1-\alpha\right)$.
\begin{proof}
$1=\left(\alpha+\left(1-\alpha\right)\right)^{N}=\sum\limits _{j=0}^{N}\bin Nj\alpha^{j}\left(1-\alpha\right)^{N-j}$
For which $j$ the term in the sum is the biggest? We want to use
derivatives, but we can only use discrete derivation, we compare to
the closest term and and check how close it it to zero: 
\begin{align*}
\bin Nj\alpha^{j}\left(1-\alpha\right)^{N-j} & \backsimeq\bin N{j+1}\alpha^{j+1}\left(1-\alpha\right)^{N-j-1}\\
\Updownarrow\\
\left(j+1\right)\left(1-\alpha\right) & \backsimeq\left(N-j\right)\alpha\Rightarrow j\backsimeq\alpha N
\end{align*}
 consequently 
\[
\frac{1}{N+1}\le\bin N{\alpha N}\alpha^{\alpha N}\left(1-\alpha\right)^{N-\alpha N}\le1
\]
 it is clear why $\bin N{\alpha N}\alpha^{\alpha N}\left(1-\alpha\right)^{N-\alpha N}\le1$
because the largest term in $1=\left(\alpha+\left(1-\alpha\right)\right)^{N}=\sum\limits _{j=0}^{N}\bin Nj\alpha^{j}\left(1-\alpha\right)^{N-j}$
is $\bin N{\alpha N}\alpha^{\alpha N}\left(1-\alpha\right)^{N-\alpha N}$
. Furthermore 1 is the maximal value and the average of a sum is less
then the largest value ($\frac{1}{N+1}$ is the average of sum valued
1 with $N+1$ elements), hence $\frac{1}{N+1}\le\bin N{\alpha N}\alpha^{\alpha N}\left(1-\alpha\right)^{N-\alpha N}$.
Now lets make the above term a bit prettier:
\begin{align*}
\frac{1}{N+1} & \le\bin N{\alpha N}\alpha^{\alpha N}\left(1-\alpha\right)^{N-\alpha N}\le1\\
 & \Updownarrow\\
\frac{1}{N+1}\cdot\frac{1}{\left(\alpha^{\alpha}\left(1-\alpha\right)^{1-\alpha}\right)^{N}} & \le\bin N{\alpha N}\le\frac{1}{\left(\alpha^{\alpha}\left(1-\alpha\right)^{1-\alpha}\right)^{N}}\\
 & \Updownarrow\\
\frac{1}{N+1}\cdot\frac{1}{2^{\log_{2}\left[\left(\alpha^{\alpha}\left(1-\alpha\right)^{1-\alpha}\right)^{N}\right]}} & \le\bin N{\alpha N}\le\frac{1}{2^{\log_{2}\left[\left(\alpha^{\alpha}\left(1-\alpha\right)^{1-\alpha}\right)^{N}\right]}}\\
 & \Updownarrow\\
\frac{1}{N+1}\cdot2^{-N\left[\alpha\log_{2}\left(\alpha\right)+\left(1-\alpha\right)\log_{2}\left(1-\alpha\right)\right]} & \le\bin N{\alpha N}\le2^{-N\left[\alpha\log_{2}\left(\alpha\right)+\left(1-\alpha\right)\log_{2}\left(1-\alpha\right)\right]}\\
 & \Updownarrow\\
2^{N\left(H\left(\alpha\right)-o\left(1\right)\right)}=\frac{2^{NH\left(\alpha\right)}}{N+1} & \le\bin N{\alpha N}\le2^{NH\left(\alpha\right)}
\end{align*}
 which means we get that $\bin N{\alpha N}$ is between $2^{NH\left(\alpha\right)}$
and $2^{NH\left(\alpha\right)-\epsilon}$ ergo $\bin N{\alpha N}=2^{1+o\left(1\right)}\cdot2^{NH\left(\alpha\right)}$,
as wanted.
\end{proof}
\end{lemma}


\section{More bounds}
\begin{rem}
we used previously $e^{tX}$ - the moment generating function. not
that the expectation of such random variable is defined as follows:
$\E{e^{tX}}=\sum_{j=0}^{\infty}\frac{t^{j}}{j!}\E{X^{j}}$
\end{rem}


\subsection{Chernoff bound}
\begin{thm}[Chernoff Bound]
\begin{flushleft}
 Let $X_{1},X_{2},\dots,X_{n}$ be independent indicator R.Vs, $X=\sum X_{i}$
$\mu=\E X$ then
\par\end{flushleft}
\begin{enumerate}
\item $\forall\delta>0$, $\p{X\ge\left(1+\delta\right)\mu}\le\left[\frac{e^{\delta}}{\left(1+\delta\right)^{1+\delta}}\right]^{\mu}$
\item $\forall0<\delta<1$, $\p{X\le\left(1-\delta\right)\mu}\le\left[\frac{e^{-\delta}}{\left(1-\delta\right)^{1-\delta}}\right]^{\mu}$
\end{enumerate}
\begin{proof}
We are proving only part 1, the second part should be similar. 

$\forall t>0$, $\p{X\ge\underbrace{\left(1+\delta\right)\cdot\mu}_{a}}=\p{e^{tX}\ge e^{ta}}\overset{markov}{\le}\frac{\E{e^{tX}}}{e^{ta}}$
. Lets work on the expectation:
\begin{align*}
\E{e^{tX}} & =\E{e^{t\sum_{i=1}^{n}X_{i}}}=\E{\prod e^{tX_{i}}}\overset{independent}{=}\prod\E{e^{tX_{i}}}
\end{align*}
Notice that if $X_{1}v,\dots X_{n}$ are independent so are $e^{tX_{1}}\dots e^{tX_{n}}$.
Because $X_{i}$ are indicators:
\[
\E{e^{tX_{i}}}=p_{i}\cdot e^{t\cdot1}+\left(1-p_{i}\right)e^{t\cdot0}=1+\underbrace{p_{i}\left(e^{t}-1\right)}_{x}\le e^{p_{i}\left(e^{t}-1\right)}
\]
 So 
\[
\E{e^{tX}}\le e^{\left(e^{t}-1\right)\cdot\sum p_{i}}
\]
 and because $\E{X_{i}}=p_{i}$ then 
\begin{align*}
\E{e^{tX}} & \le e^{\left(e^{t}-1\right)\cdot\sum p_{i}}=e^{\left(e^{t}-1\right)\sum\E{X_{i}}}=\\
 & =e^{\left(e^{t}-1\right)\E{\sum X_{i}}}=e^{\left(e^{t}-1\right)\E X}=e^{\left(e^{t}-1\right)\mu}
\end{align*}
 So $\p{X\ge a}\le e^{\left(e^{t}-1\right)\mu}$ choose $t=\ln\left(\delta+1\right)$
and we get
\begin{align*}
\p{X\ge a} & =\p{X\ge\left(\delta+1\right)\mu}\le\frac{e^{\left(e^{t}-1\right)\mu}}{e^{t\left(1+\delta\right)\mu}}=\\
 & =\frac{e^{\left(e^{\ln\left(1+\delta\right)}-1\right)\mu}}{e^{ln\left(1+\delta\right)\cdot\left(1+\delta\right)\mu}}=\frac{e^{\left(\left(1+\delta\right)-1\right)\mu}}{e^{ln\left(1+\delta\right)\cdot\left(1+\delta\right)\mu}}=\\
 & =\frac{e^{\delta\mu}}{\left(1+\delta\right)^{\left(1+\delta\right)\mu}}=\\
 & \left(\frac{e^{\delta}}{\left(1+\delta\right)^{\left(1+\delta\right)}}\right)^{\mu}
\end{align*}
\end{proof}
\end{thm}


\subsection{a brief encounter with martingales and the Azuma inequality}

According to Nati this isn't the standard way martingales are defined
but this is equivalent. 
\begin{defn}[Martingale]
 $\Omega$ is a probability space, $X:\Omega\to\R$ is a real R.V.
$\pi_{0}=\Omega$, $\pi_{0}\prec\pi_{1}\prec\pi_{2}\prec\dots$ are
partitions of $\Omega$. 

Notice that partition $i$ is a refined partition of $i-1$. For example
if $\pi_{1}$ splits $\Omega$ into two halves than $\pi_{2}$ splits
these two halves in some manner. 
\end{defn}

Now having these partitions we define the following random variables.
$X_{i}$ will be associated with partition $\pi_{i}$. $X_{0}\left(\omega\right)=\E X$.
for each $X_{i}$ we define $X_{i}$ over the part of the partition
that an atomic event emerged from. For example assume that $\pi_{1}$
is a partition of $\Omega$ into three events: $A,B,C$ if $\omega\in A$
then $X_{1}\left(\omega\right)=$ the expectation over event $A$
only. (for $B$ or $C$ this applies the same but their own expectation).
Now that we understand how these random variables are defined:
\begin{thm}[Azuma's inequality]
 Let $\Omega$ be a probability space, $X$ a real R.V on $\Omega$,
$\pi_{0}\prec\pi_{1}\prec\pi_{2}\prec\dots$ a sequence of refining
partitions and $X_{0}=\E X$, $X_{1},X_{2},\dots$ the corresponding
R.Vs. If $\forall i$ $\left|X_{i}-X_{i+1}\right|\le d_{i}$ then
$\p{\left|X_{0}-X_{n}\right|>\Delta}\le e^{-\frac{\Delta^{2}}{2\sum_{i=1}^{N}\left(d_{i}^{2}\right)}}$.
(Where $d_{i}$ is defined by us)
\end{thm}

Some intuition: We can think of each partition a different level of
a tree where the leaves are the atomic events. When we calculate $X_{i}$
for level $i$ we look from which leaf came $\omega$ and to what
node from level $i$ that leaf is a descendant. Then we look at the
expectation over all the leaves of that node and that is the value
of $X_{i}\left(\omega\right)$.


\section{The central limit theorem}
\begin{thm}[The central limit Theorem]
 Let $X_{1},X_{2},\dots$ be i.i.d R.V with expectation $\mu$ and
variance $\sigma^{2}$, then 
\[
\frac{\sum_{i=1}^{n}\left(x_{i}\right)-n\mu}{\sigma\sqrt{n}}\overset{d}{\to}\mathcal{N}\left(0,1\right)
\]
 That is $\forall a<b\in\R$ $\limn{\p{a<\frac{\sum_{i=1}^{n}\left(x_{i}\right)-n\mu}{\sigma\sqrt{n}}<b}}=\frac{1}{\sqrt{2\pi}}\int\limits _{a}^{b}e^{-\frac{x^{2}}{2}}dx$. 
\end{thm}

\begin{lyxgreyedout}
First of all note that $\E{X_{i}}=\mu$ for any $i$ and $\var{X_{i}}=\sigma^{2}$.
now when removing from each $X_{i}$ the value of the mean we get
that each $X_{i}$ have a mean 0 and now we can look at it like random
walk with $\sigma$ steps. so we will divide by $\sigma$and now we
have steps of size around $1$. so we divide by $\sqrt{n}$ which
is the expectation of random walk.%
\end{lyxgreyedout}
