\chapter{quantum computing}
\section{prelimenaries}
We are working over $\mathbb{C}$. Thus for vectors $v,w$ their inner product is defined as $<v|w>=v^*\cdot w=\sum v_i^*\cdot w_i$ ($v_i^*=a-b\cdot i$ is the complex conjugate of $v_i=a+b\cdot i$). We call a vector space over the complex numbers, with this type of inner product a Hilbert Space, and denote it by $\mathcal{H}$.

The norm of a vector $v$ is defined as $||v||=\sqrt{<v|v>}=\sqrt{\sum |v_i|^2}$. 
Also used to measure distance between vectors: $||v-w||=\sqrt{<v-w|v-w>}=||v||^2+||w||^2-<v|w>-<w|v>$.

In particular, for unit vectors $v,w$ the real part of their inner product equals $1-\frac{1}{2}||v-w||^2$

\begin{theorem}[Cauchy-Schwarz inequality]
    $|<v,w>| \leq ||v||\cdot ||w||$
\end{theorem}


\paragraph{Unitary Matrices}
A matrix $A$ is considered unitary if $A=A^-1$. The following conditions are equivalent:
\begin{enumerate}
    \item $A$ is unitary.
    \item $A$ preserves inner product: $<Av|Aw>=<v|w>$ for all $v,w$. (can be derived from moving $A=A^-1$ around).
    \item $A$ preserves norm $||Av||=||v||$ for all $v$.
    \item $||Av||=1$ if $||v||=1$
\end{enumerate}

\paragraph{Diagonalization and Singular Values}



\section{Class 1}

q-bits are basically vectors. $|0>$ is described as $\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ and $|1>$ as $\begin{pmatrix} 0 \\ 1 \end{pmatrix}$.

a q-bit $\alpha$ can have the following state$\alpha= a\cdot |0>+b\cdot |1>=(a,b)$, where $|0>$,$|1>$ are vectors (as seen before), and $a,b\in \mathbb{C}$. Additionaly, $||\alpha||=1$.

% note:
When we want to describe circular movement on the plane (using the unit circle), then, $\alpha=e^i\omega \cdot t$ for $t=0,1,2,3,\dots,\infty$.

measurements of $\alpha$ according to the basis $|0>,|1>$ gives us the results:
\begin{enumerate}
    \item $0$ in probability $|a|^2$
    \item $1$ in probability $|b|^2$
\end{enumerate}
In addition, the state after measurement is either $|0>$ or $|1>$.

Namely, measurement "collapses" the result to go either to 0 or 1.
So given $\alpha=a|0>+b|1>$, upon measurement, 
$a$ and $b$ determines the probability we'll get either $|0>$ or $|1>$.

this measurement doesn't "destroy" the value.

If we measure using poloroyde lenses, photon's polarity is a q-bit. So poloroyde lenses can help us look at the 'state'.
Given a uniform random photon, it'll either go through a poloroyde with $|0>$ polarity, or through a $|1>$ poloroyde with $0.5$ probability.

\paragraph{What happens with $2,\dots,n$ q-bits?}
Two (classical) bits can have the states $00,01,10,11$.
Thus, two q-bits can also have these states. We define them as vectors in space, where $00,01,10,11$ are translated to the vectors $|00>, |01>, |10>, |11>$.

Namely, we defined our space with words of length $2$, and a state/word of the system is defined given these vectors:
$\alpha=a \cdot |00>+b \cdot |01>+c \cdot |10>+d \cdot |11>$, where $a,b,c,d\in \mathbb{C}$ and $||\alpha||=|a|^2+|b|^2+|c|^2+|d|^2=1$.

In practice, when we want q-bits, these q-bits need to have two different "energy" levels that do not affect one another.
if the energies "leak" it is possible to still operate as long as we don't have "too much" noise.

These $2$ q-bits can run on two different systems, for instance one q-bit system in one place, and have another q-bit in a different location. 

Instead of writing $|00>$, we can also write
$|a,b>=|a>\otimes |b>$ where $\oplus$ is a tensor product.

Namely, a Hilbert space $\mathcal{H}_2\otimes \mathcal{H}_2= \mathcal{H}_4$

When we want to inspect the space, we can "collapse" it to look at what state it (like before)
\begin{enumerate}
 \item $|00>$ in probability $|a|^2$
 \item $|10>$ in probability $|b|^2$
 \item \dots
\end{enumerate}

What happens when we measure the first q-bit?
We'll get with probability $|a|^2+|b|^2$ the value $|0>$.
and the new state of the q-bits $\alpha$ is given by:
$$\alpha=\frac{a|00>+b|01>}{\sqrt{|a|^2+|b|^2}} $$

\begin{insight}
Why this formula? The full state is $\alpha = a|00> + b|01> + c|10> + d|11>$.
The first q-bit is $|0>$ in exactly the terms $|00>$ and $|01>$, so the probability of seeing $|0>$ is $|a|^2+|b|^2$ (sum the squared amplitudes of all matching terms).

After the measurement confirms the first q-bit is $|0>$, the terms $|10>$ and $|11>$
are \emph{impossible}---they are discarded.
What remains is $a|00> + b|01>$, but this no longer has unit norm.
Dividing by $\sqrt{|a|^2+|b|^2}$ \textbf{renormalizes} it so that probabilities again sum to $1$.
This is called \emph{partial collapse}: only the measured q-bit collapses; the second q-bit is still in superposition (conditioned on the outcome).
\end{insight}


if we have $n$ q-bits:

$$\alpha=\sum_{x\in\{0,1\}^n} {a_x\cdot |x>}$$ and $\sum_{x\in\{0,1\}^n} |a_x|^2=1$.


\begin{bclogo}[logo=\bcattention, couleurBarre=orange, noborder=true, couleur=white]{Exponential storage}
    Did you notice? Just to store a state, we need the sum of an exponential amount of data.
\end{bclogo}

\paragraph{operations}
Given an hilbert space $\mathcal{H}_d$ where $\dim(\mathcal{H}_d)=d$.

The allowed physical operations (which might be hard to do in a reasonal computation model) are all the linear operations that keep/reserve ("shomrot") angle and distance. Namely, all unitary operations: $\mathcal{U}:\mathcal{H}_d\to \mathcal{H}_d$ 
such that $u\cdot U^\dagger=I_d$, and $\mathcal{U}^\dagger=\bar{\mathcal{U}}^T$.
and since we use Complex numbers: 
$\mathcal{U}^\dagger=(\bar{\mathcal{U}}_{i,j})$.

Reminder: Hilbert space is a vector space over the complex numbers with an inner product.

denote two vectors $\alpha =(a_1,a_2,\dots a_d)$ and $\beta=(b_1,b_2,\dots b_d,) \in \mathcal{H}_d$
So $\alpha^\dagger=\alpha^*=(\bar{a_1},...,\bar{a_d})$. 
and $<\alpha | \beta > =\sum{a_ib_i}$



\paragraph{Possible operations on a single q-bit}
We need to define how to operate these complex operations, which are basically multiplying matrices. 
How can we do it in a single step?

\begin{enumerate}
    \item ID: the id matrix: $\begin{bmatrix}
        1 & 0  \\
        0 & 1 \\
    \end{bmatrix}$
    \item NOT: $\begin{pmatrix}
        0 & 1 \\
        1 & 0 \\
    \end{pmatrix}$
    \item Z operation (phase-flip) $Z|0> = -|0>=|0>$  $Z|1> = -|1>$ defined via $\begin{pmatrix}
        1 & 0  \\
        0 & -1 \\
    \end{pmatrix}$
    \item $H|0> = \frac{|0>+|1>}{\sqrt{2}}$ and for 1: $H|1> = \frac{|0> - |1>}{\sqrt{2}}$ (The Hadamard matrix) $H=\frac{1}{\sqrt{2}} \cdot\begin{pmatrix}
        1 &1 \\
        1 &-1 \\
    \end{pmatrix}$ 
    \item Rotation in $\theta$ (using a cos,sin matrix).
\end{enumerate}

We can create using all these functions every unitary matrix.

If we add to this collection (of operations) the "controlled not" operation matrix (a $4\times4$ matrix):
$|a,b> \overset{\text{CNOT}}{\to} |a,a+b>$, it is "easy to see" that one can build any unitary operation over $n$ q-bits.


Since all possible rotations are infinte amount of matrices, and thus an infinite amount of possible operations, we resolve to use approximations instead.
Reapplying the following operation allows us to forgo all possible rotation matrices, thus simplifying the amount of operations gates we have to define.

Thus, instead of $R_\theta$, we use powers of $T=\begin{pmatrix}
    1 &0 \\
    0 &e^{i\pi/4} \\
\end{pmatrix}$
to get closer to the actual wanted values, without needing to be able to compute rotation gates for very possible complex number.

\paragraph{EPR/CHSH Game}
Let us look at the following 2 q-bits
$|\phi>=\frac{1}{\sqrt{2}}|00>-\frac{1}{\sqrt{2}}|11>$
what happens if we apply $R_\alpha\oplus R_\beta$ to it?

We get:
$|\phi>=(R_\alpha|0>\oplus R_\beta|0>+R_\alpha|1>\oplus R_\beta|1>)/ \sqrt{2}=\cos(\alpha+\beta)(\frac{|00>-|11>}{\sqrt{2}})+\sin(\alpha+\beta)()\frac{|01>+|10>}{\sqrt{2}}$.

Why is it important?
Assume we have Alice and Bob. Alice holds the first q-bit, and bob holds the second q-bit. Then they go to different rooms. 

They play the following (co-op) game wher Alice $A$ and Bob $B$ want to win together: 

$R$ the referee chooses $x,y\in\{0,1\}$ in uniform random. $R$ gives alice $x$ and bob $y$ at the same time, withotu telling the others the value. 
Alice and Bob need to answer $a,b$ such that 
$a\oplus b=x\cdot y$.

$A$ and $B$ can win with $3/4$ probability by ignoring $x$ and $y$ and saying: $a=b=0$.
\begin{insight}
    Classically, Alice and Bob cannot communicate. Out of the 4 possible input pairs for $(x,y)$, the product $x \cdot y$ equals $0$ three times, and equals $1$ only once (when $x=1, y=1$). 
    Because they cannot know the other's input, the easiest and best strategy is to just guess the majority outcome ($0$) by constantly outputting $a=0$ and $b=0$. This guarantees they win exactly 3 out of 4 times ($75\%$).
\end{insight}


If Alice and Bob share a state $\phi$ (one of EPR states - Einstein, Pozolski, Rosen 1936), they can do better!

The strategy: $A$ given $x$ is to apply a unitary operator on the bit:

for $x=0$ applies $R_\alpha$ for $\alpha=-\frac{\pi}{16}$ and $\alpha=\frac{3\pi}{16}$ for $x=1$.
$B$ does the same for its $y$ value.

Looking at the results, $A$ and $B$ measure and the followings are the $a,b$ values:

\begin{center}
\begin{tabular}{c|c|c}
$x/y$ & 0 & 1 \\
\hline
0 & $\frac{-\pi}{8}$ & $\frac{+\pi}{8}$ \\
1 & $\frac{\pi}{8}$ & $\frac{3\pi}{8}$ \\
\end{tabular}
\end{center}

\begin{insight}
    The values in the table represent the sum of the rotation angles ($\alpha + \beta$) chosen by Alice and Bob for each possible input pair $(x,y)$. For example, if $x=0$ and $y=0$, they both use $-\pi/16$, which sums to $-\pi/8$. 
    
    After applying these rotations, Alice and Bob measure their individual qubits in the standard computational basis to get their outputs $a, b \in \{0, 1\}$. 
    The probability of their measured bits matching ($a=b$) is $\cos^2(\alpha+\beta)$, and the probability of them differing ($a \neq b$) is $\sin^2(\alpha+\beta)$. By choosing these specific angles, the probability of satisfying the winning condition $a \oplus b = x \cdot y$ comes out to exactly $\cos^2(\pm\pi/8) = \sin^2(3\pi/8) \approx 85\%$ across all four cases!
\end{insight}
The probability for $0$ is $cos^2(\pi/8)=0.85$
The probability for $1$ is $sin^2(3\pi/8)=cos^2=(\pi/8)=0.85$

Thus, in any way we have a solution better than a classical strategy.

Why does this happen? because they have a \emph{corelation} in the state over the entire system. Without passing information between them.
Namely, Quantum allow us more possibilities.

In Quantum world: the measurement changes the state and thus creates a correlation that are stronger than what we expect.
